---
title: "Histograms and Z-Scores"
author: "your name here"
date: "2026-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,        # Default width of plots in inches
  fig.height = 4,       # Default height of plots in inches
  fig.align = 'center' # Center-align plots: 'center', 'left', or 'right'
)
```

### In this lab, you will

1. Make a path/folder for this lab (e.g. `.../420/labs/lab2`) on 
your class folder and create a new R project in that folder.

2. Download the data file `2020heights.csv` from canvas ans move it to the new
folder.

3. In the following code chunk, load the `tidyverse` package and read the 
`2020heights.csv` data (you can use the `read_csv()` function).

```{r read-data}

```

4. Make a histogram of using the default bins.

```{r default-bins}

```

5. Make a histogram with too many bins (too narrow bin width)

```{r narrow-bins}

```

6. Make a histogram with too few bins (too big bin width)

```{r big-bins}

```

7. Make a histogram with ``good'' number of bins.

```{r good-bins}

```

8. Make a pretty histogram (e.g., colors for borders/fill) using the 
``Goldilocks'' bin size.

```{r pretty-bins}

```

9. Briefly summarize each of the distributions.

**ANS:** 

10. In the following code chunk, calculate the z-score of each value.

```{r z-score}

```

11. Identify the observations with the largest z-score magnitudes 
(i.e., biggest z-scores in absolute value) you can print the values using the 
following code chunk.

```{r largest-z}

```

**ANS:** 