---
title: "Sampling Distributions"
author: "your name here"
date: "2026-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,        # Default width of plots in inches
  fig.height = 4,       # Default height of plots in inches
  fig.align = 'center' # Center-align plots: 'center', 'left', or 'right'
)
```


1. Make a path/folder for this lab (e.g. `.../420/labs/lab4`) on 
your class folder and create a new R project on that path/folder and load the 
`tidyverse` package.

```{r load-tidy}

```


2. Use the `rnorm()` function to simulate data from a single experiment. The
experiment should have a sample size of 25, a mean of 100, and a standard 
deviation of 10.

```{r simulation1}

```

3. Make a plot of the frequency polygon from the data of your simulated 
experiment.

```{r plot-sim1}

```

4. Use the `replicate()` function to simulate 1000 repetitions of your 
experiment (remember that the replicate function has 2 main arguments, `n` 
which indicates the number of replications, and `expr` which takes the r 
function that will be replicated).

```{r many-exp}

```

5. Using the function `colMeans()` calculate the mean of each experiment that 
you simulated in step 4.

```{r mean-many}

```

6. Make a frequency plot of the sampling distribution of the means of all 1000
experiments. Additionally, print the standard error of this sampling 
distribution.

```{r plot-mean-many}

```

7. Repeat step 4 for this time simulate 1000 repetitions of an experiment with 
a sample size of 100, you can use the same mean and standard deviation as 
before.

```{r many-exp2}

```

8. Make a frequency plot of the sampling distribution of the means of all 1000
experiments with a sample size of 100. Additionally, print the standard error 
of this sampling distribution.

```{r plot-mean-many2}

``` 

9. Compare the sampling distribution of the mean of the two experiments you 
just simulated. What are the similarities and differences between them?

**Ans:**