---
title: "Fear of Flying"
author: "your name here"
date: "2026-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,        # Default width of plots in inches
  fig.height = 4,       # Default height of plots in inches
  fig.align = 'center' # Center-align plots: 'center', 'left', or 'right'
)
```

1. Make a path/folder for this lab (e.g. `.../420/labs/lab5`) on 
your class folder, create a new R project on that path/folder, download the data
file `FlyingVsDrivingFearData.csv` from canvas and move it to that folder.

2. In the following code chunk, load the `tidyverse` package and read the 
`FlyingVsDrivingFearData.csv` data (you can use the `read_csv()` function).

```{r load}

```

3. Use the functions `group_by()`, `summarize()`, and `mutate()` to get a 
summary tibble of the data that includes means and 95% CIs.

```{r summary}

```

4. Using the summary of the data in 3 make a plot of the mean and error bars of
each condition in the experiment (hint: you should have an example of this
plot in lab3).

```{r mean-error-plot}

```

5. Make a simulation of 1000 replications of the fear of driving condition. You 
can use the values of sample size, mean, and standard deviation (rounded to 2 
decimal places) from the data. Compute the mean of each simulated replication.

```{r simulate-driving}

```

6. Make a simulation of 1000 replications of the fear of flying condition. You 
can use the values of sample size, mean, and standard deviation (rounded to 2 
decimal places) from the data. Compute the mean of each simulated replication.

```{r simulate-flying}

```

7. Compute the sampling distribution of the difference between sample means from 
the simulated experiments.

```{r difference-sampling}

```

8. Plot the sampling distribution of the difference between means (hint: try to 
make the sampling distribution of the difference between means into a tibble 
before making the plot).

```{r plot-diff}

```

9. Perform a t-test on the difference between fear of driving and fear of 
flying. You can use the `t.test()` function which has 2 arguments, `data` which 
should be equal to the name of the variable that contains the experimental data
(the variable used when the data was originally loaded). The second argument 
is `formula` and it should be equal to `y ~ x` where y should be replaced with 
the name of the measurement and x with the name of the variable that has the 
condition.

```{r ttest}

```

10. Briefly summarize your findings. Do the values of t and p intuitively agree
with the plot of means $\pm$ 95\% CIs? Also, compare the 95\% CI on the 
difference from the t-test with your plot of the simulated differences between 
the means, do they seem to agree?

**Ans:**
