---
title: "Fear of Flying Paired Data"
author: "your name here"
date: "2026-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,        # Default width of plots in inches
  fig.height = 4,       # Default height of plots in inches
  fig.align = 'center' # Center-align plots: 'center', 'left', or 'right'
)
```

As in the previous lab, we are going to look at the fear of flying vs. driving
data. In this case though, we are going to ask for fear ratings for both modes 
of transportation from the *same people*. The idea is that if people are really 
scared of driving, they are likely really scared of flying as well. If we can
somehow get rid of this "general fear of hurling around in metal contraptions",
then we can get a more precise estimate of the difference in fear between flying
and driving.

## The data

The data are (simulated but realistic) ratings of how much each of 25 randomly
selected people fear flying and driving; each person rated both on a 10 pt. 
scale. The file contains three variables:

- id: a letter between a and y identifying the subject
- mode: "flying" or "driving"
- fear: a rating between 1 and 10.

1. Make a path/folder for this lab (e.g. `.../420/labs/lab6`) on 
your class folder, create a new R project on that path/folder, download the data
file `FearWithinData.csv` from canvas and move it to that folder.

2. In the following code chunk, load the `tidyverse` package and read the 
`FearWithinData.csv` data.

```{r load-data}

```

3. Make a frequency polygon plot of the data of fear as a function of mode of 
transportation.

```{r freq}

```

4. Make a plot of fear as a function of mode of transportation with the 
participants color-coded. You can use the `geom_jitter()` function to make
identical data points plot near to (rather than on top of) one another. Write a 
short comment on what you can see in the plot, are there ant ``general'' 
trends?

```{r fear-part}

```

**Ans:**

5. Compute the difference in fear rating for each person by subtracting their 
fear of flying from their fear of driving.

```{r calc-diff}

```

6. Make a frequency polygon plot of the difference between ratings.

```{r freq-diff}

```

7. Create a summary tibble of the difference between ratings that includes
sample size, mean, standard deviation, standard error, and the width of the 
confidence interval.

```{r summ}

```

8. Simulate 1000 replications of the difference between ratings using the sample
size, mean and standard deviation from 7. Use the `colMeans()` function to 
calculate the mean of each replication of the experiment.

```{r sim}

```

9. Make a frequency plot of the mean of each of the 1000 replications of the 
experiment that we obtained in 8. Compare this distribution with the one we 
obtained in the previous lab (`lab5.html`). What are the similarities and 
differences between the two?

```{r freq-diff-sim}

```

**Ans:**

10. Compute the t-test. For the paired t-test, we specify the paired lists of
ratings as data vectors, and we set the paired argument to TRUE. Compare the 
results with the results of the previous lab (`lab5.html`).

```{r paired-t}

```

**Ans:**

11. Write a short interpretation of the results. What do the results tell us 
about the comparison of the fear for each mode of transportation?

**Ans:**





