---
title: "First R Studio Project Lab"
author: "your name here"
date: "2026-08-13"
output: html_document
---

# Lab 1

#File structure

First, follow the instructions on the slides to create your folders and RStudio project. Then, you will complete the following:

## Part 1: Slido questions

1. Make sure that this lab1.Rmd file is in your lab1 folder.

2. Knit this file.

3. Look at the **output**, and answer the *Slido questions*.

**Code Block 1**

```{r}

# Example 1

1+1

```

**Code Block 2**

```{r}

# Example 2

a <- 1
a

# Example 3

a <- c(3,4,5)
a

# Example 4

a <- c(3,4,5)
a[3]

```

## Part 2: Viewing data

1. Download the data file `firstData.csv` from the lab website at `monaxue.github.io/420m` and put it in your `lab1` folder.

2. Load the dataset by completing the following code chunk (and remove comments if needed). Between the quotes inside the function `read_csv()`, input the name of the file (along with `.csv`).

```{r}

library(tidyverse)

#numbers_data <- read_csv("")

```
3. What name did the dataset get assigned to?

**Answer:** Your answer here.

4. What is the column name of the dataset? Use the function `head(numbers_data)` to take a peak at the top of the dataset.

```{r}



```

**Answer:** Your answer here.

5. Using using the format `dataset_name$column_name`, assign the list of numbers in the column `digits` to the variable `a`. Make sure to replace `dataset_name` and `column_name` with the actual names.

```{r}


```


6. Using the following code chunk, print the 1st, 6th and 12th values of 
the dataset using the `[]` notation.

```{r}


```

7. Introduce yourself! What are your hobbies? How do you feel about statistics and coding?

**Answer**: 

8. Knit and submit this project as an html file on Canvas.








