---
title: "Descriptive vs. inferential statistics"
author: "your name here"
date: "2026-09-08"
output: html_document
---
## Part 1: Basics of data manipulation

#### 1. In the following code block, load the two packages we always need (`tidyverse` and `ggplot2`) and import the `HandWashingData.csv` data, assigning it to the variable `washes_data`.

```{r}

```

#### 2. Use `mutate()` to find the mean of `washes`

```{r}

```

#### 3. Use `summarize()` to find the mean of `washes`

```{r}

```

#### 4. Use `group_by()` to group the data by sex, then use `summarize()` to find the mean of `washes`.

```{r}

```


## Part 2: Case study practice

### Research question

Hand washing is one of the most important things that humans can do to prevent 
the spread of many infectious diseases, such as the flu and various forms of
coronaviruses. Do females in general wash their hands more often than males?

### Methods

25 female and 25 male volunteers were given a fitbit-like device that measured 
the number of times they washed their hands. The total number of times each
volunteer washed their hands over the course of a week was measured. The 
volunteers were initially told that the devices measured “activity”, and were 
only told that hand washing was being measured after the experiment. 


### Load packages and import data

```{r}

```

### Descriptive statistics

#### 1. What are the 3 types of descriptive plots?

**Answer:**

#### 2. Create all three types of descriptive plots. Make sure create separate sub-plots for each sex.

```{r}

```


#### 3. What are the statistics to include in a descriptive summary table? 

**Answer:**


#### 4. Create a descriptive summary table and assign it to the variable `summary_table`. Then print it.

```{r}

```

#### 5. Interpret your descriptive plots. How are the plots alike/different? Based on this plot alone, can you infer if females in general wash their hands more than males?

**Answer:** 

### Inferential statistics

#### 1. Calculate the lower and upper confidence intervals for both groups. Hint: use the summary table from above.

```{r}

```

#### 2. Create a mean plot with 95% CIs.

```{r}

```

#### 3. Why do we need inferential statistics? What does it allow us to conclude that descriptive statistics alone cannot?

**Answer**