# Week 5 Lab
**Create your lab4 folder and project**

## Review question

```{=html}

<iframe src=https://wall.sli.do/event/3oY6Sijg8wvh8dq8k65iQf/?section=7c2e35e5-0092-4380-a5b1-e2828ec7cd06 width=100%, height=600></iframe>

```

# Sampling distributions

## rnorm(n, mean, sd)

::: card-list

- "r" stands for random
- "norm" stands for normal distribution
- This function samples from a normal distribution randomly. 
- Let's take see this in action.

:::

## Now you try!

::: {.card-list .nonincremental}

- Run the `rnorm(n, mean, sd)` function with a mean of 100 and an sd of 10, with a sample size of 1.

```{=html}

<iframe src=https://wall.sli.do/event/3oY6Sijg8wvh8dq8k65iQf/?section=7c2e35e5-0092-4380-a5b1-e2828ec7cd06 width=100%, height=600></iframe>

```

:::


## replicate()

::: card-list
- We can run each experiment multiple times.
- replicate(), creates a matrix of results, where each column is a single experiment.

:::

. . .

````{verbatim}
```r

matrix <- replicate(3, rnorm(5, mean = 100, sd = 10))
matrix

```
          [,1]      [,2]      [,3]
[1,] 122.85353  98.96323  99.39199
[2,]  85.13403  85.24940  93.66550
[3,]  99.45157 102.84034 119.71288
[4,] 111.03297 109.18319  84.86919
[5,]  93.45378 109.78161 104.45262

````

. . .

**Let's see this in action**



## Why do we want to simulate with replicate()?
::: card-list
- Can we just take a larger sample size?
- It allows us to calculate summary statistics for sub-samples.
- We can learn about the relationship between these summary statistics for smaller samples, and the "true" population.

:::

## Means of means and standard errors

**Let's take a look at the simulation**

## Key points

::: card-list
- A distribution of means is called the sampling distribution
- The standard deviation of the means (sampling distribution) is the standard error
- The standard error is much smaller than the standard deviation

:::

## Now you try!
::: card-list
- `samples_5 <- replicate(5, rnorm(n = 10, mean = 100, sd = 10))`
- `sampling_dist <- colMeans(5_samples)`
- `SE <- sd(sampling_dist)`

:::

:::: scrollable

## Deliverables


::: {.card-list .nonincremental}

1. From a normal distribution with `mean = 100`, and `sd = 10`, draw a sample of 25:
    - Hint: use `sample <- rnorm(n=..., mean=..., sd=...)`, and fill in the ... with values that make sense.
    - What is the mean and standard deviation?
    - Plot the sample using `geom_freqpoly()`
    - Hint: ggplot requires a dataframe. You can convert a sample to dataframes by using `tibble(x=sample)`
1. Create a sampling distribution of 1000, where each sample contains 25 samples from a normal distribution with `mean = 100` and `sd = 10`
    - Hint: use `samples_1000 <- replicate(N=..., rnorm(...))`
    - What is the mean and standard error?
    - Create a vector called `sampling_dist` by using `colMeans(samples_1000)`
    - Convert this matrix into a a dataframe for plotting (using `tibble()`)
    - Plot the sampling distribution using `geom_freqpoly()`
1. Repeat the two steps above for a sample size of 100 (also using 1000 samples for the second part)
1. Compare the two **sample** distributions (about the same width)
1. Compare the two **sampling** distributions (n=100 is narrower)
1. Show me your RMarkdown PDF when you are done.

:::

::::


```{r}

library(tidyverse)

library(ggplot2)

sample_25 <- tibble(x=rnorm(25, 100, 10))

sample_100 <- tibble(x=rnorm(100, 100, 10))

sampling_25 <- tibble(x=colMeans(replicate(1000, rnorm(25, 100, 10))))

sampling_100 <- tibble(x=colMeans(replicate(1000, rnorm(100, 100, 10))))


```

##

::: panel-tabset

### Sample n=25
```{r}

ggplot(sample_25, aes(x=x)) + geom_freqpoly()
```

### Sample n=100
```{r}
ggplot(sample_100, aes(x=x)) + geom_freqpoly()
```

### Sampling n=25

```{r}

ggplot(sampling_25, aes(x=x)) + geom_freqpoly()

```

### Sampling n=100

```{r}

ggplot(sampling_100, aes(x=x)) + geom_freqpoly()

```

:::