---
title: "Histograms and Z-Scores"
author: "your name here"
date: "2026-09-02"
output: html_document
---

## Part 1: Load packages and import data

In the following code chunk, do the following:

1. Load the package called `tidyverse`
2. Load the package called `ggplot2`
3. Import `sample_heights.csv` and assign it to the variable `heights_table`
4. Use the `head()` function to take a peak at your data table

```{r}

```

5. What table name was the imported data assigned to?

**Answer:** Your answer here.

6. What are column name(s) of the dataset?

**Answer:** 

## Part 2: Plotting histograms

1. Make a histogram of all of the heights data. This is the *true* distribution of heights. 

Fill in the `...` with the following, and replace `column_name` and `table_name` with the correct names. Remember to uncomment the code if necessary.

- aes(x=column_name)
- table_name
- geom_histogram()


```{r}

# ggplot(..., ...) + ...

```

2. Interpret and describe the distribution. What's the shape? Is there any skew? 

**Answer:** 

3. Now lets add another layer to our histogram plot, facet_wrap(). 

3a. In the following code block, replace the `...` with the histogram code from above.

Remember to uncomment the code if necessary.

```{r}
# ... + facet_wrap(~RandomSample)
```

3b. What does facet_wrap(~RandomSample) do?

**Answer**: 

3c. What is the relationship between the smaller histogram and the first large histogram?

**Answer:** 

4. Describe the differences you see between the 4 histograms. Specifically, describe differences in shape/skew and width.

**Answer:** 

5. Why might the very fine details of a histogram, like jaggedness or small differences in skew/width, NOT be worth interpreting in a case study? *Hint:* The keyword here is "random".

**Answer:** 

6. Can you draw definitive conclusions about the *true mean* (in the first histogram) based on any given RandomSample histogram? Why or why not?

**Answer:** 

## Part 3: Transforming the histogram

In part 3, we will transform Heights so it has a mean of 0 and an SD of 1.

### A. Shift the entire histogram left to have a mean of 0.

1. Take a look at the first histogram. What is the mean of the Heights?

**Answer:** 

2. In order to shift all the data points of Heights down a constant value, what operation do you need to do to the data? Addition, subtraction, multiplication, or division? 

**Answer:** 


3. For the mean to be at 0, what constant value do you need to use?

**Answer:** 


4. In the following code chunk, shift the data to the left to have a mean of 0. Replace the `...` with the mathematical expression that uses `Heights` and the constant value you identified.


```{r}

# heights_table <- heights_table %>% mutate(ShiftedHeights = ...)

```

5. Use the `head()` to take a look at heights_table. What does the `mutate()` function do?

```{r}

```

**Answer**: 

6. Plot the ShiftedHeights as a histogram

```{r}

```

### B. Make the histogram narrower to have an SD of 1.

7. In order to narrow the ShiftedHeights histogram by a constant factor, what operation do you need to do to the data? Addition, subtraction, multiplication, or division? 

*Hint*: What's an operation that makes both positive and negative numbers closer to zero without changing their sign?

**Answer:** 

8. The SD of the histogram is 2. What constant factor should you use to reduce the SD of histogram from 2 to 1?

**Answer:** 

9. In the following code chunk, make the data narrower to have an SD of 1. Replace the `...` with the mathematical expression that uses ShiftedHeights and the constant factor you identified.

```{r}

# heights_table <- heights_table %>% mutate(TransformedHeights = ...)

```

10. Plot TransformedHeights as a histogram

```{r}

```

11. In A and B, we transformed Heights to have a mean of 0 and a SD of 1. There is a name for this particular transformation. What is the name for this score?

*Hint:* You should have learned about this from the online modules. 

**Answer:** 

12. Write the full equation for this score using the following variables: `mean`, `SD`, and `x` (the value of each data point). Make sure to use parentheses properly!

**Answer:** 

13. Once you have completed all steps and checked your answers, knit the file and submit your html file output to Canvas.