---
title: "Sampling Distributions and CIs"
author: "your name here"
date: "2026-08-15"
output: html_document
---

## Part 1: single simulation experiment

#### 1. Load the two packages that we need in every class.

```{r}

```

#### 2. Now let's take a **r**andom sample from a **norm**al distribution (`rnorm()`) as a single experiment. The experiment should have a sample size of 25, a true population mean of 100, and a true standard deviation of 10. Use the `rnorm()` function and assign the simulated data to the variable `single_experiment`.

```{r}

```

#### 3. Uncomment the code below and run the code block. You ***do not*** need to understand how the code works here. I have just converted `single_experiment` into a usable data table format. Observe the output and pay attention to the column names.

```{r}

#single_experiment_table <- tibble(data = single_experiment)

#head(single_experiment_table)

```

#### 4. Make a histogram from the data of your simulated experiment.

```{r}


```

## Part 2: 100 simulation experiments

#### 1. Use the `replicate()` function to simulate 100 repetitions of your experiment (remember that the replicate function has 2 main arguments, `n` which indicates the number of replications, and `expr` which takes the r function that will be replicated).

```{r}


```

#### 2. Uncomment the code below and run the code block. You ***do not*** need to understand how the code works. I have transformed `more_experiments` data into a usable data table format for you. Take a look at the resulting table and pay attention to its column names.

```{r}

#colnames(more_experiments) <- 1:100

#more_experiments_table <- more_experiments %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="data") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

#head(more_experiments_table)

```

#### 3. Create a summary table of `more_experiments_table`, grouping by experiment. Assign it to the variable `summary_table`. Don't forget to print it.

```{r}


```

#### 4. What does the sampling distribution plot? And is this value stored in `more_experiments_table` or `summary_table`?

**Answer:** 

#### 5. Make a histogram of the sampling distribution, using the table you identified above as the data table in ggplot().

```{r}


```

#### 6. Calculate the confidence interval for each group using summary_table and assigning the results to CIs (exactly the same as what we did last week). Then print CIs.

```{r}


```

#### 7. Now create a mean/CI plot for each experiment (exactly the same as what we did last week). Then assign the plot to the variable CIplot and show the plot by printing it.

```{r}


```

#### 8. Uncomment the code below and run it. You don't have the understand the code, but I basically just added some colors to your original CI plot so things are easier to see.

```{r}

#CIplot + geom_hline(yintercept=100, color="blue") + geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI, color = (LowerCI > 100 | UpperCI < 100))) +  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red"),labels = c("Yes", "No"),name = "Overlap with 100")

```

#### 9. How many experiments have CIs that do NOT overlap with the population mean? What percent of the exerpiments is that?

**Answer:**


#### 10. The SEs you have calculated in summary_table is just an estimated SE. The true SE is actually the sd  of the sampling distribution of the means. Use `mutate()` to add a new variable to the summary_table called `TrueSE`, and calculate it by taking the sd of the means.

```{r}


```

#### 11. Fill in the blanks below by replacing the bolded text (but keep your answers bolded).

**number** is free response.

Use the word bank to fill in **answer**:
- probability
- raw values
- true mean
- confidence
- sample means


The population mean is **number**. We simulated **number** experiments, where each experiment had a sample size of **number**. 

The sample distribution is created by plotting the histogram of **answer** of a single experiment, whereas the sampling distribution is created by plotting the histogram of **answer** across all experiments. 

The 95% confidence interval is calculated such that in 100 experiments, about **number** of them would contain the **answer**. Confidence intervals do not tell us anything about **answer**, but it tells us about the **answer** that we have.

#### 12. Write interpretations for the mean/CI plot from last week's lab. 

**Answer**: 
