---
title: "Case study 1"
author: "your name here"
date: "enter the date here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,        # Default width of plots in inches
  fig.height = 4,       # Default height of plots in inches
  fig.align = 'center'  # Center-align plots: 'center', 'left', or 'right'
)
```

## Instructions
Read the background provided below and complete each deliverable. You may work with a partner on this assignment, but each student must submit their own set of deliverables. If you choose to work with a partner you must go to Canvas and set up your group on the `People` tab. If you get stuck, try looking back at the code used in previous lab assignments.

## Background
You work as a research assistant in a developmental psychology lab at UT. You’ve recently been assigned to work with one of the lab’s graduate students on a large-scale cross-cultural study, where children at many sites across the world completed a series of experimental tasks. The graduate student is making a poster to show off their preliminary work and would like your help with the descriptive analysis portion.

You will focus on just one task from the larger battery, where children watched an adult build a tower out of special blocks and were subsequently asked to build their own tower. The adult’s tower had 10 distinctive features and each child’s tower was given an imitation score (0-10) based on how many of those features they incorporated into their own tower. Each child’s age was also recorded.

## Deliverables

### 1. In your own words, describe the goal of this analysis. [2 points]

**ANS:**

### 2. Read in the data and load any necessary packages. [1 point]

```{r read-data}

```

### 3. Identify how many children participated at each field site. [2 points]

```{r sample-size}

```

### 4. Report the mean age of the participants for each field site. [2 points]

```{r mean-age}

```

### 5. Produce histograms that display the raw imitation scores for each field site. [3 points]
(Hint: facet_wrap may be useful)

```{r histograms}

```

### 6. Caculate the mean imitation score per fieldsite along with 95% confidence intervals. [2 points]

```{r mean-cis}

```

### 7. Graph the mean imitation score per fieldsite (including 95% confidence intervals). [3 points]

```{r mean-cis-graph}

```

### 8. Interpret the results in the context of the experiment (what should we take away from this data?). [3 points]
**ANS:**

### 9. Thoroughly comment all of your code above so that your collaborators can follow along with your work. [2 points]

