---
title: "First R Studio Project Lab"
author: "your name here"
date: "2026-08-13"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

### In this lab, you will

1. Make a path/folder for this lab (e.g. `home/thisSemester/420/labs/lab1`) on 
your class folder.

2. Create a new RStudio project in this folder.

3. Download the data file `firstData.csv` from Canvas and put it in the folder.

4. Import the data into RStudio. You can use the `Import Dataset` button found 
at the top of the environment tab.

5. In the following code chunk, copy and paste the code shown in your console 
that appears after you import the data.

```{r load-data, eval=FALSE}

```

6. Using the following code chunk print the first, sixth and twelve values on 
the data-set (you should use the `[` `]` notation)

```{r print-values}

```

7. Knit and submit this project as an html file, you can use the button that 
appears under the open file name in Rstudio.


