---
title: "Boxplots and Confidence Intervals"
author: "your name here"
date: "2026-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,        # Default width of plots in inches
  fig.height = 4,       # Default height of plots in inches
  fig.align = 'center' # Center-align plots: 'center', 'left', or 'right'
)
```

## Research question for this lab: 
Hand washing is one of the most important things that humans can do to prevent 
the spread of many infectious diseases, such as the flu and various forms of
coronaviruses. Do females in general wash their hands more often than males?

## Methods:
25 female and 25 male volunteers were given a fitbit-like device that measured 
the number of times they washed their hands. The total number of times each
volunteer washed their hands over the course of a week was measured. The 
volunteers were initially told that the devices measured “activity”, and were 
only told that hand washing was being measured after the experiment. 

### In this lab, you will

1. Make a path/folder for this lab (e.g. `.../420/labs/lab3`) on 
your class folder and create a new R project on that path/folder.

2. Download the data file `HandWashingData.csv` from Canvas and add it to that 
new folder.

3. In the following code chunk, load the `tidyverse` package and read the 
`HandWashingData.csv` data (you can use the `read_csv()` function).

```{r read-data}

```

4. Make a histogram of the number of `washes` by `sex` with a ``good'' binwidth.
Describe these distributions (*i.e.* how are they alike/different? Do they 
suggest any differences between the sexes?).

```{r histogram}

```

**Ans:** 

5. Make a frequency polygon of the number of `washes` by `sex`. What are the 
similarities/differences between the frequency polygon and the histogram?

```{r polygon}

```

**Ans:**

6. Make a boxplot of the number of `washes` by `sex`. Between these boxplots and
the previous summary plots, which one do you find easier to interpret and why?
When might we prefer to use one type vs another?

```{r box-plot}

```

**Ans:**

7. Create a table that contains the sample size, mean, standard deviation, 
standard error of the mean, and the width of the confidence interval of the 
number of `washes` by`sex` (result should only have 2 rows). Print the result
using the following code chunk.

```{r table}

```

8. Make a plot that displays the mean and confidence interval of the mean number 
of `washes` by `sex` (**hint:** you can use the functions `geom_point()` and 
`geom_errorbar()`).

```{r mean-ci-plot}

```

9. Write a short interpretation of the results of this experiment. What should 
we conclude about the hand washing behavior between the sexes? Should we think
they are the same-ish, or different? How can we justify our conclusion?

**Ans:**