---
title: "Descriptive vs. inferential statistics"
---

## Announcements

- Quiz next week!
- Work on your practice quizzes, they will be very similar to the actual quiz.
- Bring a handwritten cheat sheet (letter size) for extra credit.
- While the quiz is on Canvas, you must be in person to take it.
- The first half of lab will be a short, normal lab. You will have the second half of lab to complete the quiz.
- The quiz is timed and will be 40 minutes long.

## Overview
<ol class="menu"><ol>

## What are the first three steps for every lab?

1. Create lab folder
2. Create R project
3. Download and move Rmd and CSV files into lab folder from **lab website**

# Data manipulation {data-stack-name="Data manipulation"}

What you do to the data before you plot it

- Essential operators and functions
    - Pipes: `%>%`
    - `mutate()`
    - `summarize()`
    - `group_by()`


## Pipes: `%>%` {#pipes}

- Applies functions to data tables
- The following are equivalent
    - `head(table_name)`
    - `table_name %>% head()`
- Why use pipes? 
    - They allow you to chain multiple functions together
    - `table_name %>% mutate(...) %>% head()`

## `mutate()` function {#mutate}

- Adds or changes columns in a data table
    - `table_name %>% mutate(new_column = ...)`
- Example
    - `table_name %>% mutate(Mean = mean(washes))`

##

- Go to Part 1 of your `lab3.Rmd` file.
- Try to use `... %>% mutate(Mean = mean(washes))` on your dataset.
- ❓ What did `mutate(Mean = mean(washes))` do?
    - It calculated the mean of washes and replaced every value in the washes column with that mean.
    - **It created a new column called Mean, and in each row, it calculated the average of the entire column of washes. (correct)**
    - It created a new column called Mean, and for each row, it calculated the average of that row.


<!--

::: iframe

https://wall.sli.do/event/svJsdsn6gh1fzGkA76YVCv?section=9a2f41e7-317b-4de0-b355-fd79899b3503&integration=shared-present-mode&utm_source=slidoadmin

:::

-->

## `summarize()` function {#summarize}

- `table_name %>% summarize(Mean=mean(washes))`
- ❓ Compare `summarize()` with `mutate()`. How do they differ?
    - summarize() deletes rows with missing data
    - summarize() adds up all the rows so only one is left
    - **summarize() aggregates the values and removes redundant rows (correct)**

. . .

<!--

::: iframe

https://wall.sli.do/event/svJsdsn6gh1fzGkA76YVCv?section=9a2f41e7-317b-4de0-b355-fd79899b3503&integration=shared-present-mode&utm_source=slidoadmin

:::

-->

<!-- 

-->

##  group_by() {#group-by}

- We can use pipes to add the `group_by()` before using summarize.
- `table_name %>% group_by(sex) %>% summarize(Mean=mean(washes))`
- You can group by multiple groups, like so: `group_by(column1, column2, column3)`
- ❓ Compare `summarize()` with `group_by()` to the to the previous version without it. What's the difference?
    - **summarize() alone produces one summary for the full dataset, while group_by() followed by summarize() produces a separate summary for each group. (correct)**
    - group_by() sorts the dataset, and summarize() then reports the first value in each group.
    - group_by() removes duplicate rows before summarize() calculates the mean.

<!--

. . .

::: iframe

https://wall.sli.do/event/svJsdsn6gh1fzGkA76YVCv?section=9a2f41e7-317b-4de0-b355-fd79899b3503&integration=shared-present-mode&utm_source=slidoadmin

:::

-->



## Summary table {#sum_table}


- When we used `group_by()` with `summarize()`, we created a summary table.
- Generally, we have the form:

. . .

```r
summary_table <- table_name %>% group_by(...) %>% summarize(...)
```

```{r}
library(tidyverse)

washes_data <- read_csv("data/HandWashingData.csv")

summary_table <- washes_data %>% group_by(sex) %>% summarize(SampleSize = n(), Mean = mean(washes), SD = sd(washes), SE = SD/sqrt(SampleSize))

summary_table
```

# Descriptive vs. inferential statistics {data-stack-name="Descriptive vs inferential statistics"}

## Descriptive vs. inferential statistics

::::: columns

:::: column

- **Descriptive**
    - Visualize the ***general*** shape, center, and spread of the ***sample***
    - "See the forest, but not the trees"

::::

:::: column


- **Inferential**
    - Make inferences about the true population means

::::
:::::


# Descriptive statistics

- Larger trends in the data, without making claims about the true population values.
- "Ambiguous thinking"
    - Shape: Is it normal? Symmentrical? Skewed (rarer than you think!)? Bimodal?
    - Center ~ ***Sample*** mean
    - Spread ~ ***Sample*** SD

## "Ambiguous thinking"


```{r}

library(tidyverse)

set.seed(1)

n=50
x = c()
experiment = c()

for (i in 1:3){
    x = c(x, rbinom(n=n, size = 20, prob = 0.5))
    experiment <- c(experiment, rep(i, times=n))
}

df <- tibble(`Performance Score`=x, experiment=experiment)

ggplot(df, aes(x=`Performance Score`)) + geom_histogram() + facet_wrap(~experiment)

```

## Interpreting descriptive statistics {#descriptive-interpret}

1. Shape (skew)
    - All three samples are somewhat bell shaped. There is no clear skew in any of the plots.
1. Spread
    - All three samples have similar similar widths,. None are clearly wider than the other.
1. Center (optional)
    - In all three samples, the score is centered at around 10.
    - Note: Performance ***score***, not just performance.

## ❓ Which interpretation is the best? 

:::::: panel-tabset

## Histograms

```{r}

library(tidyverse)

set.seed(3)

N=25
x = c()
experiment = c()
Mean = c(5, 7)
SD = c(2, 3)


for (i in 1:2){
    x = c(x, rnorm(n=N, mean=Mean[i], sd=SD[i]))
    experiment <- c(experiment, rep(i, times=N))
}


df <- tibble(`Memory Score`=x, experiment=experiment)

ggplot(df, aes(x=`Memory Score`)) + geom_histogram(binwidth=2) + facet_wrap(~experiment)

```

## Answer options

::: nonincremental

1. **Both plots are bell shaped, with the second plot being slightly more skewed. The first plot is slightly narrower than the second. The score is centered at around 5 for the first plot, whereas it is centered at around 7 for the second plot. (correct)**
2. The second plot is significantly more skewed than the first plot. It is also significantly wider. Both plots are bell shaped. Memory levels for the second group is higher, with a mean of 7, than for the first group, with a mean of 5.
3. Both plots are rather bell shaped, though the second plot may be more skewed and is slightly wider than the first plot. The memory levels for the second group is higher than for the first group, at about 7, whereas the first group has a lower memory level, at about 5.
4. **Both plots are bell shaped. The second plot is slightly wider and slightly more skewed than the first plot. (correct)**

:::

## Explanations

- 1 includes talking about the shape, spread, and center, without making claims about the popultion.
- 2 is wrong because it uses the word "signficant", which implies a population level conclusion.
- 3 is wrong because it makes a strong statement about a group difference. Without specifying that the difference is in the the score, not just memory levels, it doesn't specify that it is only in these samples. 
- 4 avoids any issues with interpreting the "center" by not talking about it at all. Remember that in this class, describing the "center" is optional for descriptive plots.



<!--

::::: columns

:::: column

::::

:::: column

::: iframe

https://wall.sli.do/event/svJsdsn6gh1fzGkA76YVCv?section=9a2f41e7-317b-4de0-b355-fd79899b3503&integration=shared-present-mode&utm_source=slidoadmin

:::

::::

:::::

-->

::::::


# Inferential statistics

- What can we infer about the population from the sample?

##

- Guessing game: I have take a random sample from either Population A or Population B. Which population does this random sample come from?
    - "Population" = The true distribution

<!-- Just do a raise hand poll. who thinks it comes from distribution A? Who thinks it comes form distribution B? -->
. . . 

```{r}

library(patchwork)

set.seed(1)

population <-  function(mean, sd, xlim, sample_size){

    density <- round(dnorm(x=xlim, mean= mean, sd=sd)*sample_size)
    X <- numeric(0)

    for (i in 1:length(xlim)){
    X <- c(X, rep(xlim[i], times = density[i]))
    }

    X

}

A <- population(mean = 64, sd = 4, xlim = 20:110, sample_size = 100)
B <- population(mean = 68, sd = 4, xlim = 20:110, sample_size = 100)


combined <- tibble(X = c(A, B), Population = c(rep("A", times=100), rep("B", times=100)))

pop <- ggplot(combined, aes(x=X, fill=Population)) + geom_histogram(position = "identity", binwidth=1, alpha = 0.8) + xlim(c(52, 80)) + ylim(c(0,10))


samples <- tibble(`Sample 2` = rnorm(mean=64, sd=4, n= 8), `Sample 1` = rnorm(mean=68, sd=4, n= 8))


sample1 <- ggplot(samples, aes(x=`Sample 1`)) + geom_histogram(binwidth=1) + xlim(c(52, 80)) + ylim(c(0,10))

pop + sample1

```




##

I took a second random sample from either Population A or Population B. Which distribution does this random sample come from?

```{r}

sample2 <- ggplot(samples, aes(x=`Sample 2`)) + geom_histogram() + xlim(c(52, 80)) + ylim(c(0,10))

pop + sample2

```

<!-- - Actual answer: An answer that doesn't make as much sense. -->

## Statistical tests must be used

- It can be hard to tell which population the sample came from!
- Instead of asking, "do we know for sure which distribution the sample comes from?", can we instead ask "what is the probability that the sample comes from A or B?"
    - Kind of. There isn't a way to exactly calculate the probability, but we can still capture this intution with the idea of "confidence".
- We must use some sort of *test*, like confidence intervals or t-tests, to infer something about the true population (more on CIs next week)

# Doing statistics in your Case Studies {data-stack-name="Doing statistics in your Case Studies"}

How to do descriptive and inferential statistics in your case studies?

## Descriptive statistcis

:::::: panel-tabset

## What to include

In case studies, descriptive statistics always includes the following components

::: nonincremental
1. Summary table
1. Plot: histogram, frequency polygon, or boxplot
1. Interpretations
:::


## Summary table

::: nonincremental

- A table that shows sample size, mean, standard deviation and standard error
    - `summary_table <- table_name %>% group_by(...) %>% summarize(SampleSize = n(), Mean = ..., SD = ..., SE = ...)`
- To calculate the sample size of each group, just set SampleSize to the function `n()` (nothing needs to go in the parentheses)
- To calculate SE, use the function `SD/sqrt(SampleSize)`

:::

```{r}
summary_table <- washes_data %>% group_by(sex) %>% summarize(SampleSize = n(), Mean = mean(washes), SD = sd(washes), SE = SD/sqrt(SampleSize))

summary_table
```



## Histogram/frequency polygon{#histogram-freqpoly}

::: nonincremental

- Counts the number of individuals (y-axis) for the variable on the x-axis.

:::

::::: columns
:::: {.column .nonincremental}

**Histogram**

::: nonincremental

- `ggplot(..., aes(x=...)) + geom_histogram()`
- ggplot calculates the counts, so only x is needed.

:::

```{r}
data <- read_csv("data/HandWashingData.csv")
ggplot(data, aes(x=washes)) + geom_histogram()
```
::::

:::: {.column .nonincremental}

**Frequency polygon**    

::: nonincremental

- `ggplot(..., aes(x=...)) + geom_freqpoly()`
- This is the continuous version of the histogram

:::

```{r}
ggplot(data, aes(x=washes)) + geom_freqpoly()
```
::::
:::::

## Boxplot {#boxplot}

::: nonincremental

- `ggplot(..., aes(x=..., y=...)) + geom_boxplot()`
- Note: needs both x and y variables, where x is the grouping variable.

:::

```{r}

ggplot(data, aes(x=sex, y=washes)) + geom_boxplot()

```


## Interpretations

::: nonincremental
- Important for your case study!
- Include comparison of rough shape (bell shaped or skewed), width, and optionally center (if done right!)
- Please review the [example interpretations on the previous slides](#descriptive-interpret)
:::
::::::



## Inferential statistics

:::::: panel-tabset

## What to include

::: nonincremental


- Calculate confidence intervals or conduct t-tests
- CI plot
- Interpretations
:::

## Calculate CI

::: nonincremental


- Mean $\pm$ 1.96*SE for 95% CI
- In R, calculate using summary table
    - `CIs <- summary_table %>% mutate(LowerCI = Mean - 1.96*SE, UpperCI = Mean + 1.96*SE)`
:::
## CI plot {#ci-plot}

::: nonincremental

- `ggplot(CIs, aes(x=grouping_variable, y=Mean)) + geom_errorbar(aes(ymin = LowerCI, ymax=UpperCI))`
- `grouping_variable`: Make sure to replace `grouping_variable` with the actual correct variable.
:::

- If we are looking at difference in washes between the sexes, what variable should go in place of the `grouping_variable`? Hint: print out your CI table at find the column name that makes the most sense.
    - Answer: `sex`



```{r}
data <- read_csv("data/HandWashingData.csv")

summary_table <- data %>% group_by(sex) %>% summarize(Mean = mean(washes), SD = sd(washes), SE = SD/sqrt(n()), LowerCI= mean(washes) - 1.96*SE, UpperCI = mean(washes) + 1.96*SE)

ggplot(summary_table, aes(x=sex, y=Mean)) + geom_errorbar(aes(ymin = LowerCI, ymax=UpperCI))

```

## Interpretations

We will go over this next week.

::::::


# Cheatsheet {data-stack-name="Cheatsheet"}

## Cheat sheet for deliverables

All the code you need are on the slides. Remember: try to always manually type the code instead of copying and pasting.

:::::: panel-tabset

## Loading packages and importing data

::: nonincremental

- `library()`: loads packages
- `read_csv("name-of-file.csv")`: imports data from `name-of-file.csv`. Always remembering to assign it to a variable with `<-`

:::

## plots with ggplot()

::: nonincremental

- Basic usage
    - $\underbrace{\text{ggplot(table_name, aes(x=...y=...))}}_{\text{Info layer}} + \underbrace{\text{geom_...()}}_{\text{Plot layer}} + \underbrace{\text{facet_wrap(~...)}}_{\text{Grouping layer}}$
- Plot layers
    - `geom_histogram()`: histogram.
    - `geom_freqpoly()`: frequency polygon.
    - `geom_boxplot()`: boxplot.
    - `geom_errorbar()`: "error" bars or CI bars
:::

## Data manipuation and summary tables

::: nonincremental
- `%>%`: an operator that allows you to change together functions.
- `group_by()`: groups data by a column that you specify.
- `summarize()`: summarizes data across the dataset or group.
- `mutate()`: adds or changes a column.
:::

:::::::