---
title: "7: t-tests"
---

## {.center}
::: {.card-list-center}
- Go to <https://monaxue.github.io/420m> and open up this presentation. You will be answering Slido questions directly in the slides.
:::

# Part 1: Plot data

##

::: {.card-list .nonincremental}

1. Download from Canvas: `flyingVsDrivingFearData.csv`
1. Group the data appropriately
1. Create a frequency polygon
1. Print a summary table with:
    - Counts
    - Means
    - Standard deviations
    - Standard errors
    - 95% confidence intervals
5. Plot means with CIs

:::

## Slido 1

::: iframe

https://app.sli.do/event/pr7zvfTyuotunzsNJfTEii/embed/polls/7b8e3923-bb92-4b22-a3ec-54d565e5c649

:::


# Part 2: Simulating data 

## {.center}

::: card-list

- First, let us make the following assumptions:
    - Fear scores for both driving and flying are normally distributed
    - Driving has a true population mean of 2.84 and standard deviation of 1.028.
    - Flying has a true population mean of 5.20 and standard deviation of 0.957.

:::

## Slido 1.5

::: iframe

https://app.sli.do/event/pr7zvfTyuotunzsNJfTEii/embed/polls/c5bb62dd-a391-476e-a887-de5c97c0e99e

:::

##

::: {.card-list .nonincremental}

1. Using the assumed mean and standard deviation of `driving`, draw 25 samples from a normal distribution. Assign this to the variable `driving_sim_data`.
1. Replicate the above "experiment" 1000 times
1. Repeat steps 1 and 2 for `flying`.
1. Create sampling distributions for `driving` and `flying`.
    - Use colMeans() to take the column means of both matrices (from `replicate()`) and assign them to the variables `driving_means` and `flying_means`.
1. Compute the difference between means using `diff_between_means <-  flying_means - driving_means`
1. Convert `diff_between_means` into a dataframe using `tibble()` and assign this to `meanDiffsTibble`
1. Plot `meanDiffsTibble` as a frequency polygon.

:::

## Slido 2

::: iframe

https://app.sli.do/event/pr7zvfTyuotunzsNJfTEii/embed/polls/134173ee-3dd8-440e-bb48-c7e00e5b9bf8

:::

# Part 3: t-test

The function for conducting a t-test is `t.test()`. Use it like this:

````{verbatim}
```{r}

t.test(data = data, fear ~ mode)

```
````

```{r}

library(tidyverse)

data <- read_csv("data/FlyingVsDrivingFearData.csv")

t.test(data = data, fear ~ mode)

```

You should get an output like this.

## Slido 3

::: {.card-list .nonincremental}

- In the plot of CIs from Part 1, we compared the mean of driving with the mean of flying. From the `t.test()` output, can you tell what two values we are comparing?

:::

::: iframe

https://app.sli.do/event/pr7zvfTyuotunzsNJfTEii/embed/polls/142b155d-47bb-437d-bfaf-243fd03043b6

:::


## Interpret the results

::: {.card-list .nonincremental}

- Copy and paste the following into your R Markdown file, and fill in the blanks.
    - You can click on "Copy to Clipboard" on the top right corner.

:::

```{verbatim}

# Interpretation of results

The 95% confidence intervals (CIs) for fear of driving and fear of flying are __ and __, respectively. This means that we can be 95% ___ that ________. Looking at the plot of the means with CIs, I can see that the CIs of driving and flying __. This means we can be fairly __ that the true means are __.

Furthermore, the sampling distribution of the difference is centered on __, and the majority of the distribution falls between __ and __. If people didn’t fear flying more than driving, the difference should be __, which is not included in the sampling distribution. This suggests that it is unlikely that the true difference is __.

A t-test on the data confirms the above intuition. The confidence interval generated from the t-test is __. This means that we are 95% __ that ________. The range of the confidence interval is __ to the range of the sampling distribution of the difference. The p-value is __, which means that the probability of obtainting __ if our __ hypothesis is true is low. Therefore, we can conclude that ________.

```


## Summary of deliverables

::: card-list

1. Frequency polygons of the data distributions
1. Plots of the means with 95% CIs
1. A frequency polygon of the simulated sampling distribution of the difference
1. The output of the t-test
1. Interpretation of results

:::


# Slido questions and answers


# Slido 1

##

::: {.card-list .nonincremental}

- Are the frequency polynomials **literally** normally distributed?

    1. Yes 
    2. No

:::

##

::: card-list
    
- **Correct answer:** No

- **Explanation:** A normal distribution is a specific function that has exact values. While this function might describe the general shape of real world phenomenon, it does not perfectly reflect real world mechanisms for generating data. One property of a normal distribution is that the tails go on to infinity. This is not a property of most real world phenomenon. Real world data never perfectly match hypothetical functions that we define.

:::

##

::: {.card-list .nonincremental}

- Do you think samples are drawn from a population that approximates a normal distribution?
    1. Likely
    2. Unlikely

:::

##

::: card-list

- **Correct answer:** Likely
- **Explanation:** The general shape of each frequency polygon appears to have a mean with more or less even spread around it. They are not really skewed (notice that any extension of lines are actually at `count = 0`). This seems reasonable because one expects that there is a central tendency that most people agree upon, and deviations from the mean comes from random "errors" that are not systematically biased, like ambiguity of the rating scale, variations in personality, mood, etc. Generally, if a research question concerns means, you can expect what gets measured will be more or less normally distributed.

:::

##

::: {.card-list .nonincremental}

- What can you conclude about the differences between the two frequency polygons?

    1. People have higher levels of fear for flying than for driving. The peaks of the frequency polygons are offset quite noticeably with very little overlap, suggesting that the difference likely generalizes to the true population as well as future samples.

    1. In this particular dataset, there is a difference in mean ratings between the driving a flying conditions. However, we cannot conclude if this difference is generalizable to the true population or any future samples.

    1. We cannot form any conclusions from this plot alone. Even though the peaks look different, we have not plotted the confidence intervals. Therefore, we cannot be sure about the difference in means for this experiment, and by extension, the true population.

:::

##

::: card-list

- **Correct answer:** 2
- **Explanation:** The frequency polygons plot the raw data, and we can describe what we see in the raw data. However, we cannot make inferences about the likely true population differences without doing further analyses, like calculating the confidence interval or conducting a t-test. When we say that "people have different fear levels for driving and flying", we are making a general statement about people, and, therefore, the true population.

:::


# Slido 2

##

::: {.card-list .nonincremental}

- Write a plot title that clearly describes the final plot
    - **Example answer:** Frequency polygon of the differences between the mean fear of flying and mean fear of driving over 1000 simulated experiments
- What is the value of the mean of the differences between means?
    - ~ 2.3
- Where is the bulk of the distribution? (i.e., the range of the differences in means)
    - ~ 1.75 to 3

:::

# Slido 3

## 

::: {.card-list .nonincremental}

1. In the t-test, which two values are being compared? Which is the null hypothesis?

**Answer:** The difference between the means and 0, which is the null hypothesis.

1. How does the CI from the t-test compare to the bulk of the sampling distribution of differences?

**Answer**: They are fairly similar.

:::


