# Week 9
Download `salsLargeN.csv` and `salsSmallN.csv`


## Goal for today

::: card-list

- Today, we will be seeing if there is a difference in annual salary between UT grads and Texas A&M grads.

- We will look at two studies:
    1. Large sample size (`salsLargeN.csv`) 
    2. Small sample size (`salsSmallN.csv`)

- It should be noted that these datasets are fake.

:::

## 

**Since we are looking at a difference between means, what plots/analyses should we do in R?**

::: {.card-list .nonincremental}

- Frequency polygon
- t-test
- Summary table
- Means plot with 95% confidence interval

Answer: All of the above.

:::

## t-test review

::: {.card-list .nonincremental}

- `data$salary outputs a vector of salaries
- `data$school == "UT"` outputs a vector of `TRUE` and `FALSE`.
- `data$salary[data$school == "UT"]` outputs the data where `data$school == "UT"` is true (AKA the data for all UT grads)
- Therefore, `data$[data$school == "UT"]$salary` outputs the salary of all UT grads.
:::

##

:::: {.columns}

::: {.column width="30%"}

**`data`**

||school|salary|
|---|---|---|
|1|TAMU|70892|
|2|UT|71871|
|3|TAMU|73517|
|4|UT|74159|
|5|TAMU|73058|


:::


::: {.column width="10%"}

$\rightarrow$

:::

::: {.column width="20%"}

**`data$salary`**

||salary|
|---|---|
|1|70892|
|2|71871|
|3|73517|
|4|74159|
|5|73058|

:::

::: {.column width="10%"}

$\rightarrow$

:::


::: {.column width="20%"}

**`data$school == "UT"`**

||school|
|---|---|
|1|FALSE|
|2|TRUE|
|3|FALSE|
|4|TRUE|
|5|FALSE|

:::

::: {.column width="10%"}

$\rightarrow$

:::


::: {.column width = "45%"}

**`data$salary[data$school == "UT"]`**

||salary|
|---|---|
|2|71871|
|4|74159|

:::

::::

## Slido

In t.test, we want `t.test(UT_sals,TAMU_sals)`, where `UT_sals` is the salary of all UT grads, and `TAMU_sals` is the salary of TAMU grads. What should the code look like?

```r
UT_sals <- data$salary[data$school == "UT"]
TAMU_sals <- data$salary[data$school == "TAMU"]

t.test(UT_sals,TAMU_sals)

```
**Or**

```r

t.test(data$salary[data$school=="UT"], data$salary[data$school=="TAMU"])

```

## Create the plots, summary table, and do a t-test for both datasets

## Large sample

```{r}

library(tidyverse)
library(ggplot2)

data <- read_csv("data/salsLargeN.csv")

hist <- ggplot(data, aes(x=salary, color=school)) + geom_freqpoly()

UT_sals <- data$salary[data$school == "UT"]
TAMU_sals <- data$salary[data$school == "TAMU"]

t_test <- t.test(TAMU_sals,UT_sals)

summary <- data %>% group_by(school) %>% summarize(Count=n(), Mean = mean(salary), SD = sd(salary), SE=SD/sqrt(Count), CI = 1.96*SE)



mean_plot <- ggplot(summary, aes(x=school, y=Mean)) + geom_point() + geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI))

```

:::: panel-tabset

## Frequency polygon

```{r}
hist
```

## Summary table
```{r}

knitr::kable(summary)


```

## Mean plot

```{r}
mean_plot
```

## t-test

```{r}

print(t_test)

```

<!--

## Code

```r

data <- read_csv("data/salsLargeN.csv")

hist <- ggplot(data, aes(x=salary, color=school)) + geom_freqpoly()

UT_sals <- data$salary[data$school == "UT"]
TAMU_sals <- data$salary[data$school == "TAMU"]

t_test <- t.test(TAMU_sals,UT_sals)

summary <- data %>% group_by(school) %>% summarize(Count=n(), Mean = mean(salary), SD = sd(salary), SE=SD/sqrt(Count), CI = 1.96*SE)

mean_plot <- ggplot(summary, aes(x=school, y=Mean)) + geom_point() + geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI))

hist
mean_plot
knitr::kable(summary)
print(t_test)

```

-->

::::

## Small sample

```{r}
data <- read_csv("data/salsSmallN.csv")

hist <- ggplot(data, aes(x=salary, color=school)) + geom_freqpoly(binwidth=1000)

UT_sals <- data$salary[data$school == "UT"]
TAMU_sals <- data$salary[data$school == "TAMU"]

t_test <- t.test(TAMU_sals, UT_sals)

summary <- data %>% group_by(school) %>% summarize(Count=n(), Mean = mean(salary), SD = sd(salary), SE=SD/sqrt(Count), CI = 1.96*SE)

mean_plot <- ggplot(summary, aes(x=school, y=Mean)) + geom_point() + geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI))

```

:::: panel-tabset

## Frequency polygon

```{r}
hist
```

## Summary table
```{r}

knitr::kable(summary)


```

## Mean plot

```{r}
mean_plot
```

## t-test

```{r}

print(t_test)

```

<!--

## Code

```r
data <- read_csv("data/salsSmallN.csv")

hist <- ggplot(data, aes(x=salary, color=school)) + geom_freqpoly(binwidth=1000)

UT_sals <- data$salary[data$school == "UT"]
TAMU_sals <- data$salary[data$school == "TAMU"]

t_test <- t.test(TAMU_sals, UT_sals)

summary <- data %>% group_by(school) %>% summarize(Count=n(), Mean = mean(salary), SD = sd(salary), SE=SD/sqrt(Count), CI = 1.96*SE)

mean_plot <- ggplot(summary, aes(x=school, y=Mean)) + geom_point() + geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI))

hist
mean_plot
knitr::kable(summary)
print(t_test)

```

-->

::::

## Slido

::: iframe

https://wall.sli.do/event/2TAfoM6CFtGQE4tc5xhNYU/?section=a3f79d34-ac1c-479b-ba52-dd6a19e127bd

:::

<!--

## For the large dataset:

::: card-list
- Are the frequency polygons similar or different? 
    - Similar
- What is the 95% confidence interval of the difference? 
    - [11.87212, 36.63092]
- What is the t and p-value from the t-test?
    - t = 3.8397, p-value = 0.0001233
- Is the p-value larger or smaller than 0.05?
    - Smaller
- What seems "weird" about these results? Specifically, compare the frequency polygon to the p-value.
    - The p-value is small even though the frequency polygons look similar.

:::

## For the small dataset:

::: card-list
- Are the frequency polygons similar or different?
    - Different
- What is the 95% confidence interval of the difference?
    - [-914.0476, 962.5507]
- What is the t and p-value from the t-test?
    - t = 0.054301, p-value = 0.9573
- Is the p-value larger or smaller than 0.05?
    - Larger
- What seems "weird" about these results? Specifically, compare the frequency polygon to the p-value.
    - The p-value is large even though the frequency polygons look different. 

:::

-->

## What's weird?

::: card-list

**If we use a p-value of 0.05 as our significance threshold...**

- For the large dataset
    - The p-value is smaller than 0.05, so there is a significant difference.
    - But the frequency polygons look very similar!
    - So is this difference big enough to be an important difference?
- For the small dataset
    - The p-value is larger than 0.05, so we cannot reject the null hypothesis.
    - The frequency polygons look different, but since it's a small sample, it could be do to noise.

- So how can we tell if there is an important enough difference?

:::

## Effect sizes

::: card-list

- The effect size is how many standard deviations the means are apart. 

- It is a measure of signal strength, so it does not depend on sample size.

- (mean salary of TAMU - mean salary of UT) / SD of TAMU

- If you named your summary table `summary`, you can use summary[[row #,column #]] to get the values you need.

:::

. . .

```r
effect_size <- (summary[[1,3]] - summary[[2,3]]) / summary[[1,4]]
effect_size

```

```{r}
effect_size <- (summary[[1,3]] - summary[[2,3]]) / summary[[1,4]]
effect_size

```

We get 2 100ths of a standard deviation - pretty small

## Intuitive version of effect size

::: card-list

- If we select a UT grad and a TAMU grad at random, what are the chances that the UT grad makes more?
- If that comes out to around 50:50, then there probably isn’t a meaningful difference between salaries

- Plucking 1 salary from a group We can snatch one salary from a group using the `sample()` function:
    - `sample(data$salary[data$school == "UT"], size=1)`
- Comparing 2 people's salaries
    - `sample(data$salary[data$school == "UT"], size=1) > sample(data$salary[data$school == "TAMU"], size=1)`
    - If output `TRUE`, than means a single UT sample has a higher salary than a single TAMU sample.
- Then simulate this 100 times using replicate, and sum the output to count the total number of instances that UT samples are higher. Divide by 100 to get the percent.

:::

    
```r

ut_wins <- replicate(100, 
    sample(data$salary[data$school == "UT"], size=1) > 
        sample(data$salary[data$school == "TAMU"], size=1))

sum(ut_wins)/100

```
```{r}
ut_wins <- replicate(100, 
    sample(data$salary[data$school == "UT"], size=1) > 
        sample(data$salary[data$school == "TAMU"], size=1))

sum(ut_wins)/100

```



## Deliverables

::: {.card-list .nonincremental}

**For both datasets**

- Frequency polygons of the data
- A summary table of the results
- The output of a t-test
- The effect size
- The proportion or percentage of UT grads that make more than A&M grads
- Interpretation of results using the following template:

:::

```{verbatim}

The small and large sample studies actually produced [similar/different] means and standard deviations for the salaries of UT and A&M grads.

In terms of t and p-values, the two studies yielded [similar/different] results, where the large sample had a[small/large] p-value, and the small sample had a [small/large] p-value.

The effects sizes showed [a small/large] difference in salary, indicating that the differences in salary between schools are [notable/not notable], despite the [small/large] p-values for the [large/small] sample study.

```


