# Week 12
**Download `depAndSocMed.csv` from Canvas.**


## What ggplot code would you use to reproduce this plot?

```{r}

library(tidyverse)
library(ggplot2)

data <- read_csv("data/depAndSocMed.csv")

ggplot(data, aes(x=addic, y=dep)) + geom_point()

```

<!-- **Answer:** `ggplot(raw, aes(x=addic, y=dep)) + geom_point()` -->

##

::: iframe

https://wall.sli.do/event/uEK2jHiAiKEvwPckneSgMt/?section=d3e4e3b6-a11b-40e7-80e9-e9091034cc4c

:::

## New ggplot function! `geom_smooth()`


```r
ggplot(data, aes(x=addic, y=dep)) + geom_point() + geom_smooth(method = "lm")
```

```{r}

ggplot(data, aes(x=addic, y=dep)) + geom_point() + geom_smooth(method = "lm")

```

## `lm()` function
"lm" stands for "linear model". Use this function to conduct a regression analysis

```r
lm <- lm(dep ~ addic, data=data)
lm
```
```{r}
lm <- lm(dep ~ addic, data=data)
lm
```

## Which number is the slope of the regression plot?

::: iframe

https://wall.sli.do/event/uEK2jHiAiKEvwPckneSgMt/?section=d3e4e3b6-a11b-40e7-80e9-e9091034cc4c

:::

## Get more info from `lm()`

```r
summary(lm)
```

```{r}
summary(lm)
```

## Deliverables

::: card-list

- A graph of depression vs. addiction score
- A graph of depression vs. time on social media
- Regression summaries for each of those two relationships
- A brief summary of your findings

:::


