---
title: "6: p-values and t-tests"
execute:
  echo: false
  warning: false
  message: false
---

<!--
Should add an animation of what the simulation and plotting the mean differences process looks like.

-->

## ❓ Slido

::: iframe
https://wall.sli.do/event/ib94xnXY8hNMrXtzmNjiHx?section=1008f2fe-2eb5-4749-abe5-51bc37f95740&integration=shared-present-mode&utm_source=slidoadmin
:::

<!--

- What is a sampling distribution? Do you remember?
    - The distribution of individual observations in one sample collected from a population.
    - **The distribution of means from many experiments.(correct)**
    - The distribution of means from many experiments.
- Which of the following represents the definition of a confidence interval?
  - There is a 95% chance that the true mean lies within the interval
  - The CI is the range of values that the true mean is likely to be in
  - If 100 experiments are conducted, about 5 experiments will contain the true mean.
  - **5% of experiments will produce CI's that do not contain the true mean. (correct)**

-->

## Outline

<ol class="menu"><ol>

# p-values {data-stack-name="p-values"}

## Recall

- CI calculations assume we know the true mean ($\mu$)
- If actually we knew what $\mu$ is, then we can calculate probability of experiment 
- However, we don't actually know $\mu$, so we can't actually calculate probability.
- This is why we can't use probability words in our CI interpretations

## Solution: pretend we know $\mu$

- Don't make any assumptions about $\mu$
    - ❌ Cannot calculate probabilities 
    - $\rightarrow$ CIs
- Pretend that we know $\mu$
    - ✅ Can calculate probabilities
    - $\rightarrow$ p-values

# P-values

## 

::: card-list

- We will *pretend* to know $\mu$. What does this mean?
- This means that we can simulate what the sampling distribution would look like if we assume our *pretend* $\mu$ is true.
- What are good $\mu$'s to assume?

:::

## What $\mu$ to assume?

If we are interested in proving A, we cannot assume A.

::: card-list
- A circular argument: You cannot prove something if you already assume it is true.
- Examples of circular arguments:
    - I trust you because you are trustworthy
    - The scientific method works because science tells me so
    - Calibrating a compass by only using that one compass.
        - You need to use a second compass or known landmarks (like the stars).

:::

## Null hypothesis testing

- You need to make sure the pretend $\mu$ is different from the one you are actually interested in "proving".
- This is the purpose of a **null hypothesis**
    - The "opposite" hypothesis of your main hypothesis
- If our results are very different from the null hypothesis, then we falsify it.

## Plot sampling distribution of the null hypothesis

```{r}

library(tidyverse)
library(ggplot2)

area_between_norm <- function(mu, sd, alpha, minlab, meanlab, maxlab, xaxis, title){
percent <- paste0(as.character((1-alpha)*100),"%")

x <- seq(mu - 4*sd, mu + 4*sd, length.out = 1000)

df <- tibble(
  x = x,
  y = dnorm(x, mu, sd)
)

lower <- qnorm(alpha/2, mu, sd)
upper <- qnorm(1-(alpha/2), mu, sd)

mid_x <- (lower + upper) / 2
mid_y <- dnorm(mid_x, mu, sd) * 0.5

ggplot(df, aes(x=x, y=y)) +
  geom_line() +
  geom_area(
    aes(y = ifelse(x > lower & x < upper, y, 0)),
    alpha = 0.2,
    fill="steelblue"
  ) +
    labs(x = xaxis, title = title) +
  geom_vline(xintercept = c(lower, mu, upper), linetype = "dashed") + 
  scale_x_continuous(limits=c(-5,5), breaks = c(lower, mu, upper), labels = c(minlab,meanlab,maxlab)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_classic(base_size = 10) + 
   theme(
    plot.title = element_text(size = rel(1), face = "bold",
        hjust = 0.5,margin = margin(b = rel(4))),
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(linewidth = 0.5),
    axis.ticks.x = element_line(linewidth = 0.8),
    axis.ticks.length.x = unit(4, "pt"),
    axis.text.x = element_text(size = rel(2), margin = margin(t = rel(4))),
    axis.title.x = element_text(size = rel(1.5), face = "bold", margin = margin(t = rel(3))),
    plot.margin = margin(b = rel(2), r = rel(2), l=rel(2), t=rel(2)),
    panel.background = element_rect(fill = "transparent"),
  plot.background  = element_rect(fill = "transparent")) +
  annotate(
    "text",
    x = mid_x,
    y = mid_y,
    label = percent,
    size = 8
  )
}

area_between_norm(mu=0, sd=2, alpha=0.05, minlab=expression(mu-1.96 %*% SE), meanlab=expression(mu), maxlab=expression(mu+1.96 %*% SE), xaxis=expression(bar(x)), title=paste0(""))

```

Interpretation: given that the true mean is 0, the probability of getting a mean that is between $0 \pm 1.96 \times SE$ is 95%.


## {.center}

::: card-list

- Let's say in our experiment, we want to show that there is a difference between two groups. 
- The opposite hypothesis would be that there is no difference between the groups. AKA the difference = 0.
:::


##

We happen to observe the difference between two means (aka $\bar{x}_{observed}$) is $|1.96 \times SE|$.

```{r}

area_between_norm(mu=0, sd=2, alpha=0.05, minlab=expression( -bar(x)[observed]), meanlab=expression(mu), maxlab=expression(bar(x)[observed]), xaxis=expression(bar(x)), title=paste0(""))

```

❓ What's the probability that we get an experimental result that the "same" as the null hypothesis?
<!--
- **95% (correct)**
- 5%
- 2.5%
- 97.5%
-->

##

::: iframe
https://wall.sli.do/event/ib94xnXY8hNMrXtzmNjiHx?section=1008f2fe-2eb5-4749-abe5-51bc37f95740&integration=shared-present-mode&utm_source=slidoadmin
:::

<!--

::: card-list
- Then the probability of getting a mean that is between $\pm\bar{x}_{observed}$ is __?
  - Answer: 95%
:::
-->

##

```{r}

area_between_norm(mu=0, sd=2, alpha=0.05, minlab=expression( -bar(x)[observed]), meanlab=expression(mu), maxlab=expression(bar(x)[observed]), xaxis=expression(bar(x)), title=paste0(""))

```

But the null hypothesis is not the hypothesis we want to "prove". We want the opposite of the null hypothesis. What would the opposite graph of this one look like?

---

```{r}

area_outside_norm <- function(mu, sd, alpha, minlab, meanlab, maxlab, xaxis, title){
percent <- paste0(as.character((1-alpha)*100),"%")

x <- seq(mu - 4*sd, mu + 4*sd, length.out = 1000)

df <- tibble(
  x = x,
  y = dnorm(x, mu, sd)
)

lower <- qnorm(alpha/2, mu, sd)
upper <- qnorm(1-(alpha/2), mu, sd)

mid_x <- (lower + upper) / 2
mid_y <- dnorm(mid_x, mu, sd) * 0.5

ggplot(df, aes(x=x, y=y)) +
  geom_line() +
  geom_area(
    aes(y = ifelse(x < lower | x > upper, y, 0)),
    alpha = 0.2,
    fill="steelblue"
  ) +
    labs(x = xaxis, title = title) +
  geom_vline(xintercept = c(lower, mu, upper), linetype = "dashed") + 
  scale_x_continuous(limits=c(-5,5), breaks = c(lower, mu, upper), labels = c(minlab,meanlab,maxlab)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_classic(base_size = 10) + 
   theme(
    plot.title = element_text(size = rel(1), face = "bold",
        hjust = 0.5,margin = margin(b = rel(4))),
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(linewidth = 0.5),
    axis.ticks.x = element_line(linewidth = 0.8),
    axis.ticks.length.x = unit(4, "pt"),
    axis.text.x = element_text(size = rel(1), margin = margin(t = rel(4))),
    axis.title.x = element_text(size = rel(1.5), face = "bold", margin = margin(t = rel(3))),
    plot.margin = margin(b = rel(2), r = rel(2), l=rel(2), t=rel(2)),
    panel.background = element_rect(fill = "transparent"),
  plot.background  = element_rect(fill = "transparent")) +
  annotate(
    "text",
    x = mid_x,
    y = mid_y,
    label = percent,
    size = 8
  )
}


plot4 <- area_outside_norm(mu=0, sd=1.5, alpha=0.1, minlab=expression(-bar(x)[observed]), meanlab=expression(mu), maxlab=expression(bar(x)[observed]), xaxis=expression(bar(x)), title="A")

plot5 <- area_outside_norm(mu=0, sd=1.5, alpha=0.2, minlab=expression(-bar(x)[observed]), meanlab=expression(mu), maxlab=expression(bar(x)[observed]), xaxis=expression(bar(x)), title=paste0("B"))

plot6 <- area_outside_norm(mu=0, sd=1.5, alpha=0.05, minlab=expression(mu-1.96 %*% SE), meanlab=expression(mu), maxlab=expression(mu+1.96 %*% SE), xaxis=expression(bar(x)), title=paste0("C"))

area_outside_norm(mu=0, sd=2, alpha=0.05, minlab=expression(-bar(x)[observed]), meanlab=expression(mu), maxlab=expression(bar(x)[observed]), xaxis=expression(bar(x)), title="")


```


Interpretation: given that the true mean is 0, the probability of sampling our observed $\bar{x}_{observed}$ (or something more extreme) is 100%-95% = 5%.

## Definition of p-value {.center}

::: card-list-center

- The p-value is the probability of observing our results (or something more extreme) given that our null hypothesis is true.

:::

---

❓ The p-value probabilities depend on the actual observations (real data). Which of the following has the lowest p-value? 

```{r}

plot4+plot5+plot6+theme(plot.margin = margin(b = 30, r = 30, l=30, t=30),)
```

##

::: iframe

https://wall.sli.do/event/ib94xnXY8hNMrXtzmNjiHx?section=1008f2fe-2eb5-4749-abe5-51bc37f95740&integration=shared-present-mode&utm_source=slidoadmin

:::

<!--

- Answer: `A`. While `C` has the smallest area under the curve on the outside (5%), `C` is not plotting the actual observed means on the x-axis. `A` has the smallest area under the curve outside of the observed means (10%).

--->

## p-value summary

1. Assumes null hypothesis is true
2. Create a sampling distribution from the null hypothesis
3. Compare actual observed difference to the *opposite* sampling distribution (null sampling distribution)
4. The area under the curve *outside* of the observed differences line is the p-value

# p-value interpretaions {data-stack-name="p-value interpretaions"}

1. What's the probability?
2. Probability of what? 
3. Assuming the null.

## Examples

- Write an interpretation for when the p-value is 0.005
  - If the null hypothesis is true, then there is a 0.005 probability that we observe a value this extreme.
- Write an interpretation for when the p-value is 0.9
  - There is a 90% chance that we observe a value this extreme given that the null hypothesis is true.

# Deliverables {data-stack-name="Deliverables"}

## Deliverables

:::::: panel-tabset

## 1

Read in required packages and data. Name your dataset `raw_data`.

:::: code-puzzle
::: static

{{{library(tidyverse)}}}
{{{library(ggplot2)}}}

{{raw_data}} <- {{read_csv}}({{{"FlyingVsDrivingFearData.csv"}}})

head(raw_data)

:::
::::

```{r}
library(tidyverse)
library(ggplot2)

raw_data <- read_csv("data/FlyingVsDrivingFearData.csv")

head(raw_data)
```

## 2

Create a summary table of the data, comparinge fear of flying to driving. Make sure to print it.

:::: code-puzzle
::: static

{{{summary_table}}} <- {{raw_data}} %>%
 group_by({{mode}}) %>% 
 summarize(SampleSize = {{n()}}, 
 Mean={{{mean(fear)}}}, 
 SD = {{{sd(fear)}}}, 
 SE = {{SD}}/{{sqrt(SampleSize)}})

summary_table

:::
::::

```{r}

summary_table <- raw_data %>% group_by(mode) %>% summarize(SampleSize = n(), Mean=mean(fear), SD = sd(fear), SE = SD/sqrt(SampleSize))

summary_table

```

## 3

What is the difference between the two means?

**Answer**: 

## 4

Calculate CIs

:::: code-puzzle
::: static
{{CIs}} <- {{summary_table}} %>% 
{{mutate}}({{LowerCI}} = {{{Mean - 1.96*SE}}}, 
UpperCI = {{{Mean + 1.96*SE}}})

CIs
:::
::::

```{r}
CIs <- summary_table %>% mutate(LowerCI = Mean - 1.96*SE, UpperCI = Mean + 1.96*SE)

CIs

```

## 5

Make Mean/CI plots for each condition in the experiment.

:::: code-puzzle
::: static
ggplot({{CIs}}, {{{aes}}}(x={{mode}}, y={{Mean}})) + 
{{geom_errorbar}}(
  aes(ymin={{LowerCI}}, ymax={{UpperCI}}))
:::
::::

```{r}

ggplot(CIs, aes(x=mode, y=Mean)) + geom_errorbar(aes(ymin=LowerCI, ymax=UpperCI))

```

## 6

If we wanted to simulate data to test our to our actual data against it, what kind of hypothesis should we assume?

**Answer**:

## 7

Based on the type of hypothesis you identified, what should we assume as the mean difference?

**Answer**:

## 8

We happen to know that the true standard deviations for both groups is 1. If we wanted to simulate the null hypothesis, what should we make sure we choose for the means? Why?

**Answer**:


## 9

Make a simulation of 1000 replications of driving. Use a sample size of 25. Assign your results to the variable `driving`.

:::: code-puzzle
::: static

{{driving}} <- {{replicate}}(
  exp={{rnorm}}(n={{25}},mean={{0}},sd={{1}}), 
  n={{1000}})

:::
::::

```{r}

driving <- replicate(exp=rnorm(n=25,mean=1,sd=1), n=1000)

```


## 10

Make a simulation of 1000 replications of flying. Use a sample size of 25. Assign your results to the variable `flying`.

:::: code-puzzle
::: static

{{flying}} <- {{{replicate}}}(
  exp={{{rnorm}}}(n={{25}},mean={{0}},sd={{1}}), 
  n={{1000}})

:::
::::


```{r}

flying <- replicate(exp=rnorm(n=25,mean=1,sd=1), n=1000)

```

::::::

## Deliverables

:::::: panel-tabset

## 11

Uncomment the code below and run the code block. I have transformed created data table called driving_and_flying with all your simulated data.

```r

colnames(flying) <- 1:1000
colnames(driving) <- 1:1000

flying <- flying %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="flying") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

driving <- driving %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="driving") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

driving_and_flying <- merge(flying, driving, by="experiment")

head(driving_and_flying)

```

```{r}

colnames(flying) <- 1:1000
colnames(driving) <- 1:1000

flying <- flying %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="flying") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

driving <- driving %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="driving") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

driving_and_flying <- merge(flying, driving, by="experiment")

head(driving_and_flying)

```

## 12

Create the sampling distribution for both the flying and driving columns by using `summarize()` and `group_by()`. Name the new columns MeanFlying and MeanDriving. Assign this to the variable `sampling_distribution`.

*Hint: what value are is a sampling distribution composed of? You did this in week 4.*

:::: code-puzzle
::: static

{{sampling_distribution}} <- 
driving_and_flying %>% 
group_by({{experiment}}) %>% 
{{summarize}}(
  {{MeanFlying}}={{mean(flying)}}, 
  {{MeanDriving}}={{mean(driving)}})

sampling_distribution

:::
::::

```{r}

sampling_distribution <- driving_and_flying %>% group_by(experiment) %>% summarize(MeanFlying=mean(flying), MeanDriving=mean(driving))

sampling_distribution

```

## 13

Compute the sampling distribution of the difference by subtracting the mean of driving from flying using `mutate()`. Name both the column and variable name `differences`.

:::: code-puzzle
::: static

{{differences}} <- {{sampling_distribution}} %>% 
mutate(differences = {{{MeanFlying - MeanDriving}}})

differences

:::
::::

```{r}

differences <- sampling_distribution %>% mutate(differences = MeanFlying - MeanDriving)

differences

```

## 14

Plot the sampling distribution of the difference between means.

:::: code-puzzle
::: static

ggplot({{differences}}, {{{aes}}}(x={{differences}})) + 
{{{geom_histogram()}}}

:::
::::

```{r}

ggplot(differences, aes(x=differences)) + geom_histogram()

```

## 15

According the the sampling distribution of differences plot above, if we assume that null hypothesis is true, what is th probability that we sample a mean as different or larger than what you identified in question 3? How did you come to this conclusion?

**Answer**:

## 16

Now perform a t-test on the original dataset. Provide the function with the raw data, and set the score variable as a function of the grouping variable.

`t.test(data=..., score_variable ~ grouping_variable)`

:::: code-puzzle
::: static

t.test(data = {{raw_data}}, {{fear}} ~ {{mode}})

:::
::::

```{r}

t.test(data = raw_data, fear ~ mode)

```

## 17

Briefly summarize your findings. Do the p-value you identified in 10 and the one from 11 match? Do they match your conclusion from the CI plot from 4? Why?

**Answer:**

## 18

Write interpretations for the CI plot and for your p-values.

**Answer:**

::::::