# Week 8
Download `FearWithinData.csv` from Canvas

## `geom_jitter()`

::: card-list

- Gives data points a little shove for visibility

:::

. . .

```r
ggplot(data, aes(x=mode, y=fear)) + geom_jitter(width = 0.1, height = 0.1)

```

```{r}
library(tidyverse)
library(ggplot2)

data <- read_csv("data/FearWithinData.csv")

ggplot(data, aes(x=mode, y=fear)) + geom_jitter(width = 0.1, height = 0.1)


```

## 



::: card-list

- Create a a `geom_jitter()` plot with `x = mode` and `y = fear` and color by `id`.
- Your plot should look like the following. What are your observations?

:::

. . . 

```{r}

ggplot(data, aes(x=mode, y=fear, color = id)) + geom_jitter(width = 0.1, height = 0.1)

```




## Paired t-test

::::: {.card-list .nonincremental}

- Instead of computing average differences, a paired t-test uses differences for each individual.


:::: {.columns}

::: {.column width="20%"}

:::


::: {.column width="25%"}

|Name|A|B|
|----|-|-|
|Bob |2|1|
|Amy |3|2|
|Joe |4|2|
|Emma|5|1|

:::


::: {.column width="10%"}

$\rightarrow$

:::


::: {.column width="25%"}

|Name|A - B|
|----|-----|
|Bob |1|
|Amy |1|
|Joe |2|
|Emma|3|

:::

::: {.column width="20%"}

:::

::::

- To do a paired t-test in R, use the following code:
    - `t.test(data$A, data$B, paired = TRUE)` 
    - This computes `t`, `p`, and `CIs` for paired data


:::::

##

:::::: columns

::::: column

:::: card-list

- We want to calculate the differences between `flying` and `driving` for each individual using `mutate()`

- But we cannot do that with the the current dataset structure. Why?

- ::: {.iframe height="300"}
  https://wall.sli.do/event/9Bbh4F7LnAyi7JPA28eLhY/?section=06ae0710-a7c0-43cc-93d7-b74269262631
  :::

::::

:::::

::::: column

::: card-list
- What are the column names of the current dataframe?
- What column names do we need so we can use `mutate(differences = flying - driving)`. Hint:
    - |Name|A|B|A - B|
      |----|-|-|---|
      |Bob |2|1|1|
      |Amy |3|2|1|
      |Joe |4|2|2|
      |Emma|5|1|4|


:::

:::::

::::::

## `pivot_wider()`

```r
wider_data <- data %>% pivot_wider(names_from=mode, values_from=fear)

head(wider_data)

```

```{r}

wider_data <- data %>% pivot_wider(names_from=mode, values_from=fear)

head(wider_data)

```


## Deliverables

::: {.card-list .nonincremental}

- Plot the frequency polygon of the data distributions colored by `mode` (should look the same as last week's)
- Plot fear vs. mode using `geom_jitter()`, colored by subject ID
- Using `pivot_wider()`, create a wider dataframe with column names `id`, `flying`, `driving`, and assign this to `wider_data`
- Using `mutate()`, add a column named `differences` to `wider_data` that shows the difference between the fear of flying and driving for each indidivual.
- Plot the frequency polygon of the differences
- Simulate and plot
    - Print a summary table of the differences, with the count, mean, sd, se, and ci. 
    - Using the mean and sd from the summary table, simulate 1000 trials of 25 samples per trial using `replicate()` and `rnorm()`. Assign this to `diff_sim_data`
    - Take the `colMeans()` of `diff_sim_data` and then use this to create a dataframe using `tibble()`.
    - Plot the frequency polygon of the simulated means.
- Print the output of the paired t-test using 
    - `t.test(wider_data$flying, wider_data$driving, paired = TRUE)`
- A few sentences summarizing the findings and comparing these with the traditional t-test from last week.

:::

<!--

# Notes

## Simulation code and example plot

```r

wider_data <- wider_data %>% mutate(differences=flying-driving)

wider_summary <- wider_data %>% summarize(count=n(), mean=mean(differences), sd=sd(differences), ses=sd/sqrt(count), ci=1.96*ses)

diff_sim_data <- replicate(1000, rnorm(25, wider_summary$mean, wider_summary$sd))

diff_means <- colMeans(diff_sim_data)

diff_tibble <- tibble(diff_means)

ggplot(diff_tibble, aes(x=diff_means)) + geom_freqpoly(binwidth = 0.1)

```

```{r}
wider_data <- wider_data %>% mutate(differences=flying-driving)

wider_summary <- wider_data %>% summarize(count=n(), mean=mean(differences), sd=sd(differences), ses=sd/sqrt(count), ci=1.96*ses)

diff_sim_data <- replicate(1000, rnorm(25, wider_summary$mean, wider_summary$sd))

diff_means <- colMeans(diff_sim_data)

diff_tibble <- tibble(diff_means)

ggplot(diff_tibble, aes(x=diff_means)) + geom_freqpoly(binwidth = 0.1)
```

## Paired t-test results

```r
t.test(wider_data$flying, wider_data$driving, paired=TRUE)
```
```{r}

t.test(wider_data$flying, wider_data$driving, paired=TRUE)
```

::: {.card-list .nonincremental}

- The confidence interval is much smaller than from last week (which was about -1.8 to -2.9, so about twice the width). 

- The t is even bigger than last week and that the p-value is even smaller (2.2e-16 is actually the smallest value common computers can represent).

:::

## Slido answers

::: card-list

- What are the column names of the current dataframe?
    - id, mode, fear
- What column names do we need so we can use `mutate(differences = flying - driving)`.
    - id, flying, driving

:::

-->