---
title: "12: Linear Regression"
---


## Review

::: iframe

https://wall.sli.do/event/6n7e4Q8wUpUnkD2KFfhw3q

:::

<!--

- Precision is the opposite of __. The alternative hypothesis is the opposite of __. Power is the opposite of __.
    - SE, null hypothesis, p-value/alpha
 --> 

## What is the relationship between depression and addiction to social media?

We are going to simulate some data, and plot it.

## Remember this?

$$ 
y = mx + b 
$$ 

$$ 
\text{Depression} = \beta \times \text{Addiction} + \text{Error} 
$$

. . .

🖊️

## CI for regression plot

```{r}
library(tidyverse)
library(ggplot2)

data <- read_csv("data/depAndSocMed.csv")

ggplot(data, aes(x=time, y=dep)) + geom_point() + geom_smooth(method="lm")
```

Definition: 95% of experiments will produce CIs that include the true regression line.

# Deliverables

## Simple simulation of Depression vs. Addiction

:::::: panel-tabset

## Instructions

::: {.card-list .nonincremental}

1. Error ~ normally distributed, mean = 0, sd = 8, with sample size 100
2. Addiction ~ normally distributed, mean = 20, sd = 10, with sample size 100
3. Choose a $\beta$ 
4. Calculate $\text{Depression} = \beta \times \text{Social Media Addiction} + \text{Error}$
5. Plot a scatter plot of Depression vs. Addiction

:::

## Code Puzzle

:::: code-puzzle

::: static
Error <- rnorm(n = {{{100}}}, mean = {{{0}}}, sd = {{{8}}})
x <- {{{rnorm(n = 100, mean = 20, sd = 10) OR rnorm(mean = 20, sd = 10, n = 100)}}}
{{beta}} <- # choose your own number
:::

{{y}} <- beta*{{x}} + {{Error}}

df <- tibble(Depression = {{{y}}}, Addiction = {{{x}}})

ggplot({{{df}}}, {{{aes(x = Addiction, y = Depression) OR aes(y = Depression, x = Addiction)}}}) + {{geom_point()}}



::::


## Best fit line and CI

To plot best fit line and CI, use `geom_smooth()`

```r
ggplot(...) + geom_point() + geom_smooth(method = "lm")
```

```{r}

Error <- rnorm(n = 100, mean = 0, sd = 8)
x <- rnorm(n=100, mean=20, sd=10)
beta <- 0.2 # choose your own number between 0 and 1
y <- beta*x + Error
df <- tibble(Depression = y, Addiction = x)
ggplot(df, aes(x=Addiction, y=Depression)) + geom_point() + geom_smooth(method = "lm")

```

Your plot will look slightly different

## Regression summary

Find the estimated regression values using `lm()` function

"lm" stands for "linear model". Use this function to conduct a regression analysis

```r
lm <- lm(Depression ~ Addiction, data=df)
lm # gets basic info
summary(lm) # gets more info
```
```{r}
lm <- lm(Depression ~ Addiction, data=df)
lm # gets basic info
summary(lm) # gets more info
```

::::::

## Simple simulation of Depression vs. time spent on social media

:::::: panel-tabset

## Instructions

::: {.card-list .nonincremental}

- Same as before, but instead of Addiction, we now have the variable "Time"
- The mean time on social media is 6 hours. The standard deviation is 2.
- Choose your own `beta` again. Do you think time spent on social media, when someone is not addicted to it, is positively or negatively correlated with depression? The value should be between -1 and 1.

:::

## Final plot

Your plot may look different, especially if you chose a different `beta`.

```{r}

Error <- rnorm(n = 100, mean = 0, sd = 8)
x <- rnorm(n=100, mean=6, sd=2)
beta <- -0.2
y <- beta*x + Error
df <- tibble(Depression = y, Time = x)
ggplot(df, aes(x=Time, y=Depression)) + geom_point() + geom_smooth(method = "lm")

```

::::::

## Write a CI definition {.center}

In your own words, what is the definition of a CI for a regression line?

## Final deliverables checklist

::: {.normal-list .nonincremental}

- [ ] Depression vs. social media addiction score
    - [ ] Simulation code
    - [ ] A graph that includes the scatter plot, best fit line, and CI
    - [ ] Regression summary
- [ ] Depression vs. time on social media
    - [ ] Simulation code
    - [ ] A graph that includes scatter plot, best fit line, and CI
    - [ ] Regression summary
- [ ] Definition of a CI for a regression line, in your own words

:::

