---
title: "7: Paired data and t-tests"
execute:
  echo: false
  warning: false
  message: false
---

## Outline

<ol class="menu"><ol>

# Quiz 2 info {data-stack-name="Quiz 2"}


## Quiz 2
- Quiz next week!
- No lab report next week!
- Cheatsheet can only contain formulas
- Do your practice quizzes and problem sets

## Quiz 2 topics

::: nonincremental

1. Population and samples review
2. Sampling distribution review
3. p-values review
4. Skew, central tendency
5. Variability
6. Student's t
7. One-sample t
8. Pooled variance
9. Independent t and p
10. Paired t
11. Effect size

:::


# Paired t-tests {data-stack-name="Paired t-tests"}

## Jitter plot, colored by `id`

```{r}

library(tidyverse)
library(ggplot2)

raw_data <- read_csv("data/FearWithinData.csv")

```

```{r}

my_jitter <- position_jitter(width = 0.1, height = 0.1, seed = 42)

ggplot(raw_data, aes(x=mode, y=fear)) + geom_point(aes(color = id), position = my_jitter)

```
<!--
## ❓ What do you notice?

*Hint:* look at the colors!

::: iframe
https://wall.sli.do/event/seE7FxjbKBE55v5k77ud1G?section=7e94a248-0f4a-4876-b281-78a3cfca6226&integration=shared-present-mode&utm_source=slidoadmin
:::

-->

## ❓ What do you notice?

*Hint:* look at the colors!

```{r}

first_five_ids <- unique(raw_data$id)[c(1, 6, 12, 22, 25)]
raw_data$show_line <- ifelse(raw_data$id %in% first_five_ids, raw_data$id, NA)

paired_diff <- ggplot(raw_data, aes(y=fear, x= mode)) + geom_point(position = my_jitter, aes(color = id)) + geom_path(aes(group=id, color = show_line), position = my_jitter) +  scale_color_discrete(na.value = "transparent")

paired_diff

```
::: nonincremental

- The same colors are on similar levels in both groups.
- e.g., People who score lower in driving also score lower in flying

:::

## 

**Mean differences between the two groups as a whole**

<!-- Next time also show plot with geom_points are black, then show that ids turned out to be scrambled per group, so per individual trends can counter per group trends. When only looking at per group trends, the per individual trends can be all sorts of directions (aligned, misaligned, scrambled). So if per individual trends are aligned, we are more certain of the differences. -->

```{r}

raw_data <- raw_data %>% group_by(mode) %>% mutate(Mean=mean(fear))

mean_diff <- ggplot(raw_data, aes(y=fear, x= mode)) + geom_point(position = my_jitter, aes(color = id)) + geom_point(aes(y=Mean)) + geom_path(aes(y=Mean, group = 1)) 

mean_diff

```

::: nonincremental
- Less certainty
:::

##

::::: {.columns .nonincremental}
:::: {.column .nonincremental}

**Less certainty**

```{r}
mean_diff
```

- More spread
- Higher SE
- Wider CIs

::::
:::: {.column .nonincremental}

**More certainty**

```{r}
paired_diff
```

- Differences have less spread
- Lower SE
- Narrower CI

::::
:::::

##  How do we get lower SE?

Remember: SEs are the SDs of the **sampling distribution**

```{r}
source("_scripts/area_under_norm.R")

normal <- area_under_norm(mu=0, sd=1) + xlim(c(-3,3)) + ylim(c(0,1))
wider_ci <- ggplot() + geom_errorbar(aes(y = 0, x= 0, ymin=-2, ymax=2)) + ylim(c(-5,5)) + xlim(c(-1,1))

paired <- area_under_norm(mu=0, sd=0.5) + xlim(c(-3,3)) + ylim(c(0,1))
narrower_ci <- ggplot() + geom_errorbar(aes(y = 0, x= 0, ymin=-1, ymax=1)) + ylim(c(-5,5)) + xlim(c(-1,1))

```

::::: {.columns .nonincremental}
:::: {.column .nonincremental}
**Unpaired differences**

```{r}

normal

```

$\downarrow$

```{r}

wider_ci

```

Higher SE and wider CI



::::
:::: {.column .nonincremental}

**Paired differences**

```{r}

paired

```

$\downarrow$

```{r}

narrower_ci

```

Lower SE and narrower CI

::::
:::::

## Key to produce different sampling distributions

::::: columns
:::: column
### Unpaired

1. Simulate 2 sampling distributions
1. Take differences



::::
:::: column

### Paired

1. Take differences
1. Simulate single sampling distribution

::::
:::::

## Unpaired

:::::: columns
::::: column
**(1) Calculate SD**

|  Group  |     Mean    |   SD  |
| ------- | ----------- | ----- |
| driving | $\bar{x_1}$ | $s_1$ |
| flying  | $\bar{x_2}$ | $s_2$ |
:::::
::::: column

**(2) Simulate sampling distributions**

`rnorm(`$\bar{x_1}$, $s_1$`)` and `rnorm(`$\bar{x_2}$, $s_2$`)`

```{r}
flying <- ggplot() + xlim(-4, 4) +  geom_function(fun = dnorm) + xlab("Flying")
driving <- ggplot() + xlim(-4, 4) +  geom_function(fun = dnorm) + xlab("Driving")
differences <- ggplot() + xlim(-4, 4) +  geom_function(fun = dnorm) + xlab("Differences") 

flying + driving
```

:::::
::::::

:::::: columns
::::: column

**(3) Take differences**

:::: columns
::: {.column width="45%"}

```{r}
flying
```
Flying
:::

::: {.column width="10%"}
 

$-$

:::

::: {.column width="45%"}

```{r}
driving
```
Driving
:::
::::

:::::
::::: column

**(4) Get SE**

```{r}
differences + geom_vline(xintercept = 0) + geom_segment(aes(x = 0, y = 0.242, xend = 1, yend = 0.242), color = "blue")

```

:::::
::::::

## Paired

:::::: columns
::::: column
**(1) Take differences**

|  id  | flying | driving | differences |
| ---- | ------ | ------- | ----------- |
| Bob  |   2    |    1    |      1      |
| Amy  |   3    |    2    |      1      |
| ...  |  ...   |   ...   |     ...     |


:::::
::::: column

**(2) Calculate SD**

|  Group  |     Mean    |   SD  |
| ------- | ----------- | ----- |
| differences | $\bar{x_\Delta}$ | $s_\Delta$ |


:::::
::::::

:::::: columns
::::: column

**(3) Simulate sampling distribution**

`rnorm(`$\bar{x_\Delta}$, $s_\Delta$`)`

```{r}
differences
```

:::::
::::: column

**(4) Get SE**

SD of sampling distribution

```{r}
differences + geom_vline(xintercept = 0) + geom_segment(aes(x = 0, y = 0.242, xend = 1, yend = 0.242), color = "blue")

```

:::::
::::::




# Deliverables {data-stack-name="Deliverables"}


## Part 1

:::::: panel-tabset

## 1

Load the necessary packages and dataset. Assign the dataset to the variable `raw_data`. Take a peek at the data using `head()`.

:::: code-puzzle
::: static
{{{library(tidyverse)}}}
{{{library(ggplot2)}}}

raw_data <- {{{read_csv("FearWithinData.csv")}}}

head({{{raw_data}}})
:::

::::

```{r}

library(tidyverse)
library(ggplot2)

raw_data <- read_csv("data/FearWithinData.csv")
head(raw_data)

```

## 2

Make a frequency polygon plot of the data of fear as a function of mode of transportation. Use the `color` argument in `aes()`, instead of `facet_wrap()` to separate out the groups this time. In `geom_freqpoly()`, set binwidth equal to 1.

:::: code-puzzle
::: static
ggplot({{{raw_data}}}, {{{aes}}}({{{x}}} = {{fear}}, 
{{color}} = {{mode}})) + {{geom_freqpoly}}(binwidth = 1)
:::
::::

```{r}

ggplot(raw_data, aes(x=fear, color=mode)) + geom_freqpoly(binwidth=1)

```

## 3

Approximately what is the combined width of the two frequency polygons?

**Answer**:

## 4

### `geom_jitter()`

`geom_jitter()` is very similar to `geom_point()`, but it pushes the points around a bit so it's more visible. 

::::: {.columns .nonincremental}
:::: {.column .nonincremental}
`geom_point()`
```{r}

ggplot(raw_data, aes(x=mode, y=fear, color = id)) + geom_point()

```

::::

:::: {.column .nonincremental}
`geom_jitter()`

```{r}

ggplot(raw_data, aes(x=mode, y=fear, color = id)) + geom_jitter(width = 0.1, height = 0.1)

```

::::
:::::


## 5

Make a jitter plot showing all the participants' responses, where the y axis is the ratings, and the x axis is the group. Make sure the color each partipants responses.

`ggplot(..., aes(..., color = ...)) + geom_jitter(width = 0.1, height = 0.1)`

:::: code-puzzle
::: static
ggplot({{{raw_data}}}, {{{aes}}}(x = {{{mode}}}, y = {{{fear}}}, color = {{{id}}})) + 
{{{geom_jitter}}}(width = 0.1, height = 0.1)
:::
::::


```{r}

ggplot(raw_data, aes(x=mode, y=fear, color = id)) + geom_jitter(width = 0.1, height = 0.1)

```


## 6

What are the column names of the current data table?

**Answer:**

## 7

We want to compute differences between flying and driving with `mutate(differences = flying - driving)`, but currently the format of the data makes it difficult. What columns names do we need in order to compute the differences? 

*Hint:*

|  id  | flying | driving | differences |
| ---- | ------ | ------- | ----------- |
| Bob  |   2    |    1    |      1      |
| Amy  |   3    |    2    |      1      |
| Joe  |   4    |    2    |      2      |
| Emma |   5    |    1    |      4      |

**Answer:**

## 8

### `pivot_wider()`

The pivot wider function takes a column of group names and a column of scores, and creates separate columns for each group where the rows of that column are populated by the scores.

::::: {.columns .nonincremental}
:::: {.column .nonincremental}

From this:

|  id  |   mode  | fear |
| ---- | ------- | ---- |
| Bob  | flying  |  2   |
| Amy  | flying  |  3   |
| Joe  | flying  |  4   |
| Emma | flying  |  5   |
| Bob  | driving |  1   |
| Amy  | driving |  2   |
| Joe  | driving |  2   |
| Emma | driving |  1   |

::::
:::: {.column .nonincremental}

To this:

|  id  | flying | driving |
| ---- | ------ | ------- |
| Bob  |   2    |    1    |
| Amy  |   3    |    2    |
| Joe  |   4    |    2    |
| Emma |   5    |    1    |

::::

:::::


## 9

Use `pivot_wider()` to reformat `raw_data `, and assign the new table to `wider`. Take a peak with `head()`.

`pivot_wider(names_from = column_with_group_names, values_from=column_with_ratings)`

:::: code-puzzle
::: static
{{{wider}}} <- {{{raw_data}}} %>% pivot_wider(names_from = {{{mode}}}, values_from = {{{fear}}})

head({{{wider}}})
:::
::::


```{r}

wider <- raw_data %>% pivot_wider(names_from = mode, values_from=fear)
head(wider)

```


## 10

Add a new column that computes flying minus driving.

:::: code-puzzle
::: static
{{{wider}}} <- wider %>% mutate({{{differences}}} = {{{flying - driving}}})
:::
::::

```{r}

wider <- wider %>% mutate(differences = flying - driving)

```

## 11

Make a frequency plot of the difference between ratings with binwidth = 1

:::: code-puzzle
::: static
ggplot({{{wider}}}, aes({{{x}}} = {{{differences}}})) + 
geom_{{freq}}{{poly}}({{binwidth = 1}})
:::
::::

```{r}

ggplot(wider, aes(x=differences)) + geom_freqpoly(binwidth = 1)

```

## 12

What is the approximate width of this frequency polygon?

**Answer:**

::::::

## Part 2

:::::: panel-tabset

## 1

Create a summary table of the *differences* between ratings that includes sample size, mean, standard deviation, standard error.

:::: code-puzzle
::: static

summary_table <- {{{wider}}} %>% {{{summarize}}}(SampleSize = {{{n()}}}, 
Mean = {{{mean(differences)}}}, SD = {{sd(differences)}}, 
SE = {{SD}}/{{sqrt}}({{SampleSize}}))

{{{summary_table}}}


:::
::::

```{r}

summary_table <- wider %>% summarize(SampleSize = n(), Mean = mean(differences), SD = sd(differences), SE = SD/sqrt(SampleSize))

summary_table

```

## 2

Calculate the CI of differences.

:::: code-puzzle
::: static

CIs <- {{{summary_table}}} %>% {{{mutate}}}(LowerCI = {{{Mean}}} - 1.96\*{{{SE}}}, 
UpperCI = {{{Mean}}} + 1.96\*{{{SE}}})

{{{CIs}}}

:::
::::



```{r}

CIs <- summary_table %>% mutate(LowerCI = Mean - 1.96*SE, UpperCI = Mean + 1.96*SE)

CIs

```

## 3

What is the difference between the Upper and Lower CI, and is this narrower or wider than the CIs from the previous lab?

**Answer:**


## 4 

Simulate 1000 replications of the difference between ratings using the sample size and standard deviation from your summary table. Assign this to `simulated_data`.

:::: code-puzzle
::: static

simulated_differences <- {{replicate}}(n = {{1000}}, 
exp={{rnorm}}(n = {{25}}, mean = 0, sd = {{0.4899}}))

:::
::::

```{r}

simulated_differences <- replicate(exp=rnorm(n = 25, mean = 0, sd = 0.4899), n = 1000)


```
## 5 

Magic code that you don't need to worry about to transform your `simulated_differences` to a more usable format. This new table is called `simulated_differences_table`.

```r

colnames(simulated_differences) <- 1:1000

simulated_differences_table <- simulated_differences %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="differences") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

head(simulated_differences_table)

```

```{r}

colnames(simulated_differences) <- 1:1000

simulated_differences_table <- simulated_differences %>% as_tibble() %>% pivot_longer(cols = everything(), names_to ="experiment", values_to="differences") %>% mutate(experiment = as.integer(experiment)) %>% arrange(experiment)

head(simulated_differences_table)

```


## 6 

Create a summary table of `simulated_differences_table` that includes SampleSize, Mean, SD, and SE of `differences` for each experiment. Assign it to `simulation_summary`.

:::: code-puzzle
::: static

simulation_summary <- simulated_differences_table %>% 
{{{group_by}}}({{{experiment}}}) %>% 
{{{summarize}}}(SampleSize = {{{n()}}}, 
Mean = {{{mean(differences)}}}, SD = {{{sd(differences)}}}, 
SE = {{{SD/sqrt(SampleSize)}}})

{{{simulation_summary}}}

:::
::::

```{r}

simulation_summary <- simulated_differences_table %>% group_by(experiment) %>% summarize(SampleSize = n(), Mean = mean(differences), SD = sd(differences), SE = SD/sqrt(SampleSize))

simulation_summary

```

## 7 

Make a histogram of the sampling distribution using the values in `simulation_summary`.

:::: code-puzzle
::: static

{{{ggplot}}}({{{simulation_summary}}}, aes(x = {{{Mean}}})) + {{{geom_histogram()}}}

:::
::::

```{r}

ggplot(simulation_summary, aes(x = Mean)) + geom_histogram()

```

## 8 

Compare this distribution with the one we obtained in the previous lab. What are the similarities and differences between the two?

**Answer:**

## 9 

Recall that last week, we compared the actual observed mean difference in the original dataset (2.36) to the sampling distribution. If we do the same thing with today's sampling distribution, does today's null sampling distribution "theoretically" provide a larger or smaller p-value than the one from last week? Why?

**Answer:**


## 10 

Compute the t-test. For the paired t-test, we specify the paired lists of ratings as data vectors, and we set the paired argument to TRUE. Compare the p-value with the one from the previous lab. 

*Hint:* How many zeros come after the decimal point? Look at the value after "e"

`t.test(wider$column1, wider$column2, paired = TRUE)`

:::: code-puzzle
::: static

t.test({{{wider}}}\${{{driving OR flying}}}, {{{wider}}}\${{{flying OR driving}}}, paired = {{{TRUE}}})

:::
::::

```{r}

t.test(wider$driving, wider$flying, paired = TRUE)

```

**Answer:**

## 11

What is the main point of this lab? Fill in the bolded sections.

::: nonincremental
- SE is just the **(fill in)** of the sampling distribution. When we calculate paired differences, that **increases/decreases** our SD, which **increases/decreases** our SE, resulting in a **wider/narrower** CI and **smaller/larger** p-value. This means we are **more/less** certain that there is a true difference between the groups.

:::

::::::