# Week 5 Lab
**Create your lab4 folder and project**

::: notes
This week, we will plot confidence intervals, simulate (?) CIs, and go over case study expectations. But we can also do central limit theorem again and simulate using the same concept as week 3. Or maybe confidence intervals will still be covered in week 6 and won't be as important for the first case study. What about simulating the null hypothesis?

People learn better when getting personal feedback, even when you teach sometthing in class (think in dance). So maybe have a clear rubric, but do spend too much time teaching people how to write a report, but do that during the feedback session.

Go with the "try it", then "now change this" approach.

:::

Two weeks ago, in someone's lab report report, someone wrote this:

I understand where the confusion might come from...

## Review question

```{=html}

<iframe src=https://wall.sli.do/event/3oY6Sijg8wvh8dq8k65iQf/?section=7c2e35e5-0092-4380-a5b1-e2828ec7cd06 width=100%, height=600></iframe>

```

# Sampling distributions



## Simulating normal distributions with rnorm()

Because we have already demonstrated that the distribution of height becomes normally distributed via the CLT, we can now just use a function that simulates a normal distribution directly.

## rnorm(n, mean, sd)

::: card-list

- "r" stands for random
- "norm" stands for normal distribution
- This function randomly samples from a normal distribution.
- Let's take see this in action.

:::

## Slido

- Run the `rnorm(n, mean, sd)` function with a mean of 66 and an sd of 10, with a sample size of 200.

::: iframe

:::




## Central limit theorem again
Only change 2 things:
- Simulating multiple experiments, instead of each person in an experiment.
- Taking the mean instead of the sum.

```r

all_experiments <- c()

for (experiment_number in 1:100){
    sample_size <- 10 # sample size per experiment
    all_heights <- rnorm(n=sample size, mean=66, sd=10) # simulates 10 people's height
    all_experiments[experiment_number] <- mean(all_heights)
}
```

## What this means
- Instead of a distribution of individual heights, we get a distribution of mean heights (from separate experiments)
- Last time, more "causes" made the distribution more normal. This time, larger sample size makes the distribution more normal
- Last time, we simulated how a sample distribution becomes more normally distributed. This time, we simulate how a samplING distribution becomes more normally distributed.

**Let's see this in action**

## Means of means and standard errors

**Let's take a look at the simulation**

## Key points

::: card-list
- A distribution of means is called the sampling distribution
- The standard deviation of the means (sampling distribution) is the standard error
- The standard error is much smaller than the standard deviation

:::

## Now you try!
::: card-list
- `samples_5 <- replicate(5, rnorm(n = 10, mean = 100, sd = 10))`
- `sampling_dist <- colMeans(5_samples)`
- `SE <- sd(sampling_dist)`

:::

:::: scrollable

## Deliverables


::: {.card-list .nonincremental}

1. From a normal distribution with `mean = 100`, and `sd = 10`, draw a sample of 25:
    - Hint: use `sample <- rnorm(n=..., mean=..., sd=...)`, and fill in the ... with values that make sense.
    - What is the mean and standard deviation?
    - Plot the sample using `geom_freqpoly()`
    - Hint: ggplot requires a dataframe. You can convert a sample to dataframes by using `tibble(x=sample)`
1. Create a sampling distribution of 1000, where each sample contains 25 samples from a normal distribution with `mean = 100` and `sd = 10`
    - Hint: use `samples_1000 <- replicate(N=..., rnorm(...))`
    - What is the mean and standard error?
    - Create a vector called `sampling_dist` by using `colMeans(samples_1000)`
    - Convert this matrix into a a dataframe for plotting (using `tibble()`)
    - Plot the sampling distribution using `geom_freqpoly()`
1. Repeat the two steps above for a sample size of 100 (also using 1000 samples for the second part)
1. Compare the two **sample** distributions (about the same width)
1. Compare the two **sampling** distributions (n=100 is narrower)
1. Show me your RMarkdown PDF when you are done.

:::

::::


```{r}

library(tidyverse)

library(ggplot2)

sample_25 <- tibble(x=rnorm(25, 100, 10))

sample_100 <- tibble(x=rnorm(100, 100, 10))

sampling_25 <- tibble(x=colMeans(replicate(1000, rnorm(25, 100, 10))))

sampling_100 <- tibble(x=colMeans(replicate(1000, rnorm(100, 100, 10))))


```

##

::: panel-tabset

### Sample n=25
```{r}

ggplot(sample_25, aes(x=x)) + geom_freqpoly()
```

### Sample n=100
```{r}
ggplot(sample_100, aes(x=x)) + geom_freqpoly()
```

### Sampling n=25

```{r}

ggplot(sampling_25, aes(x=x)) + geom_freqpoly()

```

### Sampling n=100

```{r}

ggplot(sampling_100, aes(x=x)) + geom_freqpoly()

```

:::


