---
title: "Lab 6"
author: Your Name
output: pdf_document
---

## Goals
For this lab, we are going to look at some data showing how much people say that they fear flying vs. driving. For context, flying is much safer than driving - lifetime odds of dying by car are about 1:100, while those for dying by plane are about 1:7000 (see, e.g. https://injuryfacts.nsc.org/all-injuries/preventable-death-overview/odds-of-dying/). 

* Plot the data
  - raw data distributions
  - means with (CLT-based) 95% CIs
* Simulate (Monte Carlo style) the difference between average fear for flying vs. driving
* Compare the MC results with a traditional t-test

The fact that the simulation agrees so well with the t-test should reinforce the idea that the t-test (and all statistical tests), are mathematical "simulations" of an infinite number of replicated experiments.

## The data
The data are (simulated but realistic) ratings of how much 25 randomly selected people fear driving and how much 25 other randomly selected people fear flying, both on a 10 pt. scale. The file contains two variables: mode: "flying" or "driving", and fear: a rating between 1 and 10.

## Deliverables
Students should submit a lab report including the following:

* frequency polygons of the data distributions
* plots of the means with 95% CIs
* a frequency polygon of the simulated sampling distribution of the difference
* the output of the canned t-test
* a sentence or 3 summarizing the findings based upon the plots and comparing these with the traditional t-test

## 0. Housekeeping
* Take attendance .
* Remind them about office hours and encourage attendance as needed.
* Just generally check in and see how they are doing.
* Remind them of the next major thing due.

## 1.0 It's coding time!

### 1.1 Get the tidyverse tools and load the data
```{r, message=FALSE}
library(tidyverse)
fvd <- read_csv('FlyingVsDrivingFearData.csv')
```

### 1.2 look at frequency polygons of the data
```{r}
ggplot(fvd, aes(x = fear, color = mode)) + geom_freqpoly(binwidth = 1)
```

### 1.3 summarize the data 
```{r, message=FALSE}
fearSum <- fvd %>% 
  group_by(mode) %>% 
  summarize(
    cts = n(),
    mns = mean(fear),
    sds = sd(fear),
  ) %>% 
  mutate(ses = sds/sqrt(cts)) %>% 
  mutate(cis = 1.96*ses)
  
knitr::kable(fearSum)
```

### 1.4 plot means with CIs
```{r}
ggplot(fearSum, aes(x = mode, y = mns)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mns - cis, ymax = mns + cis), width = 0.33) +
  ylab("Fear Rating") +
  xlab("Mode of Travel")
```
Okay, well, I guess those groups are different haha. As the students make this plot, encourage them to appreciate how rare (virtually impossible) is would be for a typical flying mean to come from the distribution of driving means, and vice verse. Encourage them to look at this plot as a visual t-test.

### 1.5 Simulate the mean fear of driving across many experiments

As before, we are going to use the function `replicate()` to get data from many experiments from a population. We'll use the `rnorm()` function to make the data inside of `replicate()`. Rather than worry about indexing into the summary tibble, we'll just enter the sample size, mean, and standard deviations by hand.
```{r}
driving_sim_data <- replicate(1000, rnorm(25, mean = 2.84, sd = 1.028))
```

The `driving_sim_data` object that this creates is a 25 x 1000 matrix (table) of numbers, in which each of the 1000 columns holds the n=25 sample of data from a single simulated experiment. 

#### now compute the means for fear of driving
```{r}
# take the mean of all the samples (i.e. mean of each column)
driving_means <- colMeans(driving_sim_data)
```

### 1.6 Simulate the mean fear of flying across many experiments

Here, we'll do the same thing as above, just using the mean and sd for flying instead of driving.
```{r}
flying_sim_data <- replicate(1000, rnorm(25, mean = 5.20, sd = 0.957))
```

As above with `driving_sim_data`, we're creating a flying_sim_data object that is a 25 x 1000 matrix (table) of numbers, in which each of the 1000 columns holds the n=25 sample of data from a single simulated experiment. 

#### the **sampling** distribution for fear of flying
```{r}
# take the mean of all the samples (i.e. mean of each column)
flying_means <- colMeans(flying_sim_data)
```

### 1.7 compute the sampling distribution of the difference between means
Compute the literal differece between the means of all our simulated experiments
```{r}
diff_between_means <- driving_means - flying_means
```
### make a tibble for easier plotting 
```{r}
meanDiffsTibble <- tibble(diff_between_means)
```
### plot
```{r}
meanDiffsPlot <- meanDiffsTibble %>% 
  ggplot(aes(x = diff_between_means)) + 
  geom_freqpoly(binwidth = 0.1)

show(meanDiffsPlot)
```

So there is about a 2.3 difference in how much people fear flying vs. driving (on a 10 point scale). Notice also that the bulk of this distribution lies between about -1.75 and -3; this can be compared with the 95% CIs on the difference generated by the t-test function.

### 2.0 Look at the t-test
Compute the t-test. 

```{r}
t.test(data = fvd, fear ~ mode)
```

Well all righty then. Not that the large value of t and the teensy weensy itty bitty p-value agree with the plot of means +/- 95% CIs. Notice also that the 95% CI on the difference agrees quite will with our plot of the simulated differences between the means.

(Also, as we point out in the module, Welch gets the credit now for adjusting the df to allow for unequal variances. Poor Student.)

### 3. Cheat sheet
Here is a cheat sheet of R commands they will need.

New commands:

* `t.test()` - compute t, p, and CIs on the difference between means
* `ylab()` and `xlab()` - add axis labels to a plot

Familiar commands:

* `library()` - load a package (usually the tidyverse)
* `read_csv()` - load the data
* `ggplot()` - define data mappings for a plot
* `geom_freqpoly` - plot a frequency distribution
* `group_by()` - group data by a variable value
* `summarize()` - make a summary table
* `n()` - count the number of observations
* `mean()` - compute the arithmetic mean
* `sd()` - compute the standard deviation
* `mutate()` - add a variable to a tibble
* `print()` - print the value of a variable to the console
* `geom_point` - plot data points
* `geom_errorbar` - add error bars to a plot
* `replicate()` - do many experiments and make a table of the data
* `rnorm()` - generate normally distributed random numbers (simulated data!)
* `colMeans()` - take the mean of each column of data
* `tibble()` - make a plane R data object into a tidyverse tibble

