---
echo: false
warning: false
---

## Case Study 3

:::: panel-tabset

## Option 1

### The Stroop Effect

[Dataset for Option 1](/data/CaseStudy3_StroopEffect.csv){.filelink}

You are interested in people's ability to ignore irrelevant visual information when under time pressure. To this end, you have run an experiment in which you presented participants with color words written in various ink colors and asked them to report the ink color of each word as quickly as possible. There are two conditions in the experiment: First, a congruent condition in which the word presented is written in the same color ink, for example, the word “red” is written in red (red). Second is the incongruent condition, in which the word presented is written in different colored ink, for example, the word “red” written in blue (red). The stimuli in the study consisted of 4 colors (red, blue, yellow and green) that could be used for both the written words and the ink. Each stimulus was presented 3 times in the congruent condition and 3 times in the incongruent condition (once in each color different from the word). For each presentation of a stimulus, the experimenters recorded the time it took the participant to verbally produce the color of ink that the word was written in. So when presented with the word “red” written in blue ink (red), the participant should say “blue.” 

You have hypothesized that it takes additional effort for people to inhibit the irrelevant perceptual information from the obligatory reading of the color word, and therefore that the mean response time in the incongruent condition will be higher than in the congruent condition. You should test this hypothesis using the data provided and produce an appropriate effect size estimate. As usual, make sure you plot and discuss the raw data and relevant data summaries. Additionally, you have a collaborator who would like to perform a replication of the study. For their power analysis they want to be conservative and use the smallest effect size consistent with your data. Based on the limits of your confidence interval for the effect of congruency on reaction time, make a recommendation for the effect size they should use in their power analysis.   

## Option 2

### Memory Decline

[Dataset for Option 2](/data/CaseStudy3_MemoryDecline.csv){.filelink}

You are studying how human memory declines as we age. You have data from a longitudinal study that tested participants’ ability to remember a list of words at various points in their life: Participants were tested once at 25 years of age and again at 65 years of age. In the experiment, you presented participants with a list of 15 words that they must memorize. 24 hours later they are brought back to the lab and asked to reproduce from memory the entire list of words. If a word is recalled it is recorded as Recalled=1 and if not it is recorded as Recalled=0. 

You have hypothesized that the ability to recall long lists of items gets worse as we age, which would result in fewer words being recalled on average at age 65 than at age 25. You should test this hypothesis using the data provided and produce an appropriate effect size estimate. As usual, make sure you plot and discuss the raw data and relevant data summaries. Additionally, you have a collaborator who would like to perform a replication of the study. For their power analysis they want to be conservative and use the smallest effect size consistent with your data. Based on the limits of your confidence interval for the effect of aging on recall performance, make a recommendation for the effect size they should use in their power analysis.

## Rubric (overview)

### Overview (15 points total)

1. Formatting: Submit an html file generated by Rmd that is well formatted and organized (1 point)
1. Research Question and Hypotheses (2 points)
1. Methods Overview (2 points)
1. Results and Interpretation (7 points)
1. External sources (3 points) 

## Detailed Rubric

### Detailed Rubric

**Formatting** (1 point)

- [ ] Submit an html file generated by Rmd that is well formatted and organized (1 point)

**Research Question and Hypotheses** (2 points)

Include 1-2 sentences for each of the following:

- [ ] The research questions (there are two)
- [ ] The main hypotheses
- [ ] The null hypotheses

**Methods Overview** (2 points)

- [ ] Explain how the the research question informs your methedological choices. For instance, why did you chose to compare certain groups? 
- [ ] What is the purpose of a null hypothesis, and how does it relate to the main hypothesis?
- [ ] What plots and statistical analyses will you be conducting to answer the research question? Why did you choose them?

**Results and Interpretation** (8 points)

- [ ] At least one summary table, including sample sizes, means, SDs, SEs. (1 point)
- [ ] At least one descriptive plot (1 point)
- [ ] One inferential plot (1 point)
- [ ] A few sentences reporting key results, i.e., the means, sd, confidence intervals, p-values. (1 point)
- [ ] Interpretations (3 points)
    - [ ] What is the *main* purpose of descriptive vs. inferential statistics?
    - [ ] Include appropriate interpretations for your plots (don't overinterpret!)
    - [ ] Based on the data, what conclusions can you draw about the population?
    - [ ] Support your conclusions using the proper statistical concepts. 
        - [ ] Correctly define each statistical concept.
        - [ ] Interpret your results based on these statistical concept and its definition. Make sure to use precise language. For example, do not use terms that imply probability when that is not part of the definition of the concept.
    - [ ] Do not use the term "significant"!
 - [ ] Make the proper effect size recommendation, and explain your reasoning. See hints. (1 point)


**External sources section** (2 points)

The case studies can be completed without using any external sources. However, if you choose to use them, you must be able to justify the choice. Points will be deducted if you use content not covered in class and do not explain them in this section. 

- [ ] List every line of code and statistical concept that we have not covered in class, and explain:
    - [ ] What they do or mean
    - [ ] Why you have chosen to use this in place of what we learned in class
    - [ ] A link to the source, if it exists, otherwise a description is fine.

*If you do not use any external sources, just say "Did not use external sources."*

## Hints

### Hints

1. Is the data paired? What kind of test do you use for paired data?
2. What's the difference between the CI outputted from your t-test, vs. the CI we typically calculate in the summary table? Which one should we actually use and plot? Why? (extra credit: 1 point)
3. To calculate the effect size, instead of using the mean, use the smallest estimate consistent with our data. Based on the confidence interval, what is that value?


### `pivot_wider()` hints

Because there are multiple rows per participant per condition/Visit (condition for Stroop Effect, Visit for Memory Decline), before using `pivot_wider()`, you must first transform the dataframe into having 1 row per condition/Visit for each participant (2 rows in total for each participant). You can do this using a combination of `group_by()` and `summarize()`, and combine those multiple rows into a single row using `mean()` or `sum()`.

> You want one row per participant per condition/Visit. So which variables do you include in `group_by()`?

After transforming the dataframe, you can then use `pivot_wider()`. If your table is looking odd, double check to make sure you used the new variable you created from the previous step, instead of the old one.

Below, I show what the dataframes should look like for each option.

```{r}
library(tidyverse)
library(ggplot2)
data1 <- read_csv("data/CaseStudy3_StroopEffect.csv")
data2 <- read_csv("data/CaseStudy3_MemoryDecline.csv")

```

**Stroop Effect**

Before using `pivot_wider()`

```{r}
wider1 <- data1 %>% group_by(id, condition) %>% summarize(mean_reaction_time=mean(reaction_time))
wider1
```

After using `pivot_wider()`

```{r}
wider1 %>% pivot_wider(values_from = mean_reaction_time, names_from = condition)
```


**Memory Decline**

Before using `pivot_wider()`

```{r}
wider2 <- data2 %>% group_by(ID, Visit) %>% summarize(TotalRecalled = sum(Recalled))
wider2
```

After using `pivot_wider()`

```{r}
wider2 %>% pivot_wider(values_from = TotalRecalled, names_from = Visit)
```


### Relevant slides

- Hypothesis generation: [Lab 7](/presentations/week7.qmd)
- Descriptive vs. Inferential statistics: [Lab 4](/presentations/week4.qmd), [Lab 6](/presentations/week6.qmd)
- Plots: [Lab 4](/presentations/week4.qmd)
- Histogram interpretations: [Lab 4](/presentations/week4.qmd), [Lab 6](/presentations/week6.qmd)
- CI calculation and interpretations: [Lab 4](/presentations/week4.qmd), [Lab 6](/presentations/week6.qmd), [Lab 7](/presentations/week7.qmd), [Online simulation](https://shiny.rit.albany.edu/stat/confidence/), [Lab11](/presentations/week11.qmd)
- p-value calculation and interpretations: [Lab 7](/presentations/week7.qmd), [Lab11](/presentations/week11.qmd)
- Paired t-test: [Lab 8](/presentations/week8.qmd)
- Effect size calculation: [Lab 9](/presentations/week9.qmd)

::::