---
title: "In-Lab exercise for polynomial (linear) fitting"
---

## Goals
Today, we are going to look at how working out might affect one's life. We are going to look at some simulated (and not necessarily, but possibly realistic) data looking at how the number of hours per day somebody spends working out is related to:

  * how "in shape" people report they feel
  * resting heart rate (lower is considered more healthy)
  * productivity, as assessed by a boss or supervisor
  
If we think about this for a hot minute, it seems reasonable that as people work out more:

  * they probably feel more in shape
  * their resting heart rate probably decreases, but only up to a point
    - initially, heart rate will decrease from unhealthy to healthy levels
    - once a healthy level is reached, heart rate can't drop (much) more
  * their boss has mixed feelings
    - a little working out probably boosts productivity along with health
    - a LOT of working out might be viewed as a distraction ("Why is so-and-so at the gym when everybody else is here putting together the big project?")

There are three data sets. For each, the first variable is average hours per day spent working out. The second variable is, for each data set in turn:

* a self-report of how "in shape" each person feels on a scale of 0 to 10
* average resting heart rate in beats per minute (bpm)
* a "boss-report" productivity rating on a scale of 0 to 10

## Deliverables
You should submit a lab report including the following.

For each data set, please show a plot of 

* the relevant dependent variable (DV) vs. hrs/day working out with a 1st order polynomial (straight line) fit
* a plot of the residuals "left over" from the straight line fit
* the DV vs. hrs/day with a 2nd order polynomial (curvy line) fit
* a plot of the residuals left over from the curvy fit

There should be a sentence or two summarizing each plot. Finally, at the end, please write a short paragraph summarizing the big picture. Are all the data described by straight lines? Do all data necessarily increase or decrease overall? What information do residuals give you?

The data sets are in the files:

* shapeVsHours.csv
* bpmVsHours.csv
* prodVsHours.csv

## 0. Using this notebook
If you are already comfortable with R notebooks, forge ahead! If you are not, either look at the one of the other labs that used them (Decisions or Power), or ask your TA for help.

## 1. Load the first dataset and take a look

### first things first: load the tidyverse
```{r}
library(tidyverse)
```

Load the first dataset
```{r, message=FALSE}
inShapeData <- read_csv('./data/shapeVsHours.csv')
```

### 1.1 Fit the data with a straight line
```{r}
plt_a <- inShapeData %>% 
  ggplot(aes(x = hours, y = reported_shape)) +
  geom_point() + 
  stat_smooth(method = 'lm') +
  xlab('workout hrs/day') + 
  ylab('self report in-shape')
show(plt_a)
```


(remember, you can save this plot by right-clicking or two-fingered-tapping)

### 1.2 Plot the residuals to check the fit

Fit the data with a straight line above in order to get the residuals.
```{r}
a <- summary(lm(inShapeData$reported_shape ~ 
                  poly(inShapeData$hours, 1)))
```

Now plot the residuals - the "leftovers" after the fit.
```{r}
tib_a_res <- tibble(x = inShapeData$hours, y_res = a$residuals)
plt_a_res <- tib_a_res %>% 
  ggplot(aes(x = x, y = y_res)) +
  geom_point() + 
  stat_smooth(method = 'lm', color = 'red', fill = 'pink') +
  xlab('workout hrs/day') + 
  ylab('in-shape residuals')
show(plt_a_res)
```

Note the lack of any pattern or trend in the residuals. This indicates that our straight-line fit has captured the essence of the data except for the inevitable random error.  

Visually compare the above two plots, focusing on the left-most 10 data points to make sure you understand the relationship between these plots.

(remember, you can save this plot by right-clicking or two-fingered-tapping)

---

## 2. Load the bpm dataset and take a look
Load the heart rate data.
```{r, message=FALSE}
heartRateData <- read_csv('./data/bpmVsHours.csv')
```

### 2.1 Fit the data with a straight line
```{r}
plt_b <- heartRateData %>% 
  ggplot(aes(x = hours, y = bpm)) +
  geom_point() + 
  stat_smooth(method = 'lm') +
  xlab('workout hrs/day') + 
  ylab('self report in-shape')
show(plt_b)
```
  
  Notice that this fit overestimates the data near the middle (the data points are mostly below the line), and underestimates the data at the ends (the data points are mostly above the line). This indicates that there is a curve in the data that the straight line fit is not capturing.  
  
  We can see this more clearly by looking at the residuals...

### 2.2 Plot the residuals to check the fit

Fit the data with a straight line above in order to get the residuals.
```{r}
b <- summary(lm(heartRateData$bpm ~ 
                  poly(heartRateData$hours, 1)))
```

Let's look at the residuals.
```{r}
tib_b_res <- tibble(x = heartRateData$hours, y_res = b$residuals)
plt_b_res <- tib_b_res %>% 
  ggplot(aes(x = x, y = y_res)) +
  geom_point() + 
  stat_smooth(method = 'lm', color = 'red', fill = 'pink') +
  xlab('workout hrs/day') + 
  ylab('heart rate residuals')
show(plt_b_res)
```
  
  From this plot, we can clearly see the pattern of under- and overestimation. This fit is not doing it for us.  
  
  So let's try a fit with a curve.  
  
### 2.3 Fit the data with a curved line
Fit the data with a linear second order polynomial.  
```{r}
plt_b2 <- heartRateData %>% 
  ggplot(aes(x = hours, y = bpm)) +
  geom_point() + 
  stat_smooth(method = 'lm', formula = y ~ poly(x, 2)) +
  xlab('workout hrs/day') + 
  ylab('heart rate (bpm)')
show(plt_b2)
```
  
  That looks much better! Now let's check the residuals.  
  
### 2.4 Plot the residuals for the second order fit
Fit the data with a curved line above in order to get the residuals.
```{r}
b2 <- summary(lm(heartRateData$bpm ~ 
                  poly(heartRateData$hours, 2)))
```

Plot the residuals.
```{r}
# plot residuals for 2nd order fit 
tib_b2_res <- tibble(x = heartRateData$hours, y_res = b2$residuals)
plt_b2_res <- tib_b2_res %>% 
  ggplot(aes(x = x, y = y_res)) +
  geom_point() + 
  stat_smooth(method = 'lm', color = 'red', fill = 'pink') +
  xlab('workout hrs/day') + 
  ylab('heart rate residuals (2nd order)')
show(plt_b2_res)
```
  
  Great! Now the residuals look random, indicating that our fit has accounted for the systematic trend in the data, and only random error remains.  
  
## 3. Load the productivity dataset and take a look
Put code to load the productivity data here. You can copy and paste the code from above, and modify as needed.
```{r}

```
  
  Remember to re-label the y axes in your plots below!!!
  
### 3.1 Fit the data with a straight line
Fit the productivity data with a straight line (first order polynomial). Use the code from above, modified as necessary.
```{r}

```

### 3.2 Plot the residuals to check the fit
Do the regression.
```{r}

```
Plot the residuals from the regression to check the fit.
```{r}

```

### 3.3 Fit the data with a curved line
Plot the data with second order linear fit (a curved line).  
```{r}

```

### 3.4 Plot the residuals for the second order fit
Do the regression.
```{r}

```
Plot the residuals from the regression to check the fit.
```{r}

```

## 4. Summary
Once you've exported/saved your figures, write a few sentences summarizing the big picture. How does hrs/week of workout affect self-perceived shape, heart rate, and perceived productivity?


