---
title: "10: Precision"

---

## Goals
- What is precision?
- How does precision relate to sample size?
- How does precision relate to SD?


## Review

::: iframe

https://wall.sli.do/event/dHrUNcA6HQFhszkEDJZsyW

:::

<!--
- How do you measure width/spread of a sample distribution?
    - SD
- What's the difference between a sample distribution and a sampling distribution?
    - A sample distribution is the histogram of data from a single experiment.
    - A sampling distribution is the histogram of means from multiple experiments.
- How does SE relate to SD?
    - The SE is the SD of a sampling distribution

-->

# Precision

## Definitions
- Precision = how *narrow* a sampling distribution is
    - The narrower sampling distribution, the higher the precision
- SE is the measure of how *wide* the sampling distribution is, so precision is a measure of how *narrow* a sampling distribution is
- Therefore, precision is the opposite of SE
    - $\text{Precision} = \frac{1}{\text{SE}}$


## {.center}

|Width of a Sample Dist| Width of Sampling Dist. | Narrowness of a Sampling Dist.|
|---|---|---|
|SD|SE|Precision = 1/SE|

```{r}
library(tidyverse)
library(ggplot2)
library(patchwork)

```

<!--

##

```{r}

library(tidyverse)
library(ggplot2)
library(patchwork)

df = tibble(x= seq(-4, 4, length.out = 1000), y=dnorm(x=seq(-4, 4, length.out = 1000)))

plot1 <- ggplot(df, aes(x=x, y=y)) + geom_line() + geom_vline(xintercept=0, color="steelblue") + theme_classic() + labs(title="A: Width of Sample Dist.") + scale_y_continuous(expand = c(0, 0))
plot2 <- ggplot(df, aes(x=x, y=y)) + geom_line() + geom_vline(xintercept=0, color="steelblue") + theme_classic() + labs(title="B: Width of Sampling Dist.") + scale_y_continuous(expand = c(0, 0))
plot3 <- ggplot(df, aes(x=x, y=y)) + geom_line() + geom_vline(xintercept=0, color="steelblue") + theme_classic() + labs(title="C: Narrowness of Sampling Dist.") + scale_y_continuous(expand = c(0, 0))

plot1 + plot2 + plot3

```

-->

# Deliverables

## Deliverables Summary

::: nonincremental
- Sampling distribution plots
    1. Constant sample size and SD
    1. Varied sample size and constant SD
    1. Constant sample size and varied SD
- Plot of precision vs. sample size
- Plot of precision vs. sample standard deviation
- Interpretations
:::


## 1. Sampling distribution simulation

:::::: panel-tabset


## Code Puzzle

::::: {.code-puzzle}

::: static

SampleSize <- 49
SD <- 12
Mean <- 3
AllExperiments <- 1:5000

:::

simulated_means <- {{c()}}

::: indent
for (experiment in {{AllExperiments}}){
    
    simulated_data <- {{rnorm(n = SampleSize, mean = Mean, sd = SD)}}
    simulated_means[experiment] <- mean(simulated_data)

} 
:::

sampling_distribution <- tibble(Means={{simulated_means}})

:::: flex

::: free
plot <- ggplot({{{data=sampling_distribution OR sampling_distribution}}}, {{{aes(x = Means)}}}) + 
    geom_histogram(binwidth=0.1) + geom_vline(xintercept = 0) + xlim(c(-20,20))
show(plot)
:::

::: free

{{{summary_table <- sampling_distribution}}} %>% 
    summarize(SE = {{sd(Means)}}, Precision = {{1/SE}})

summary_table

:::

::::

:::::

<!--

## Instructions

Do the following all in one code block:

::: {.normal-list .nonincremental}

- [ ] Define variables: SampleSize, SD, Mean, AllExperiments at the top of your code block
    - [ ] `SampleSize <- 49`
    - [ ] `SD <- 12`
    - [ ] `Mean <- 3`
    - [ ] `AllExperiments <- 1:5000`
- [ ] Set up your simulation
    - [ ] Create an empty list named `simulated_means` to store simulation results
    - [ ] Create your `for loop` to repeat experiments

    ```r
    for (experiment in AllExperiments){
        ...

    }
    ```

- [ ] Inside your `for loop`
    - [ ] Sample data from a normal distribution using `rnorm(n = ..., mean = ..., sd = ...)`
    - [ ] The means of each experiment is saved to the variable `simulated_means`

- [ ] Create a dataframe named `sampling_distribution` using `tibble(Means = ...)`
- [ ] Plot a histogram of the sampling distribution using:

    ```r
        ggplot(...) + geom_histogram(binwidth=0.1) + geom_vline(xintercept = 0) + xlim(c(-20,20))
    ```

    - Notes: `geom_vline(xintercept = 0)` plots a vertical line at 0, and `xlim(c(-20,20))` makes sure the x-axis ranges from -2- to 20.
- [ ] Create a summary table with just `SE` and `Precision`
- [ ] Make sure everything on this page is in a single code block

:::
-->

## Plot

```{r}

SampleSize <- 49
SD <- 12
Mean <- 3
AllExperiments <- 1:5000
simulated_means <- c()
for (experiment in AllExperiments){
    simulated_data <- rnorm(n = SampleSize, mean = Mean, sd = SD)
    simulated_means[experiment] <- mean(simulated_data)
}
sampling_distribution <- tibble(Means=simulated_means)
summary_table <- sampling_distribution %>%
summarize(SE = sd(Means), Precision = 1/SE)
summary_table
plot <- ggplot(sampling_distribution, aes(x=Means)) +
geom_histogram(binwidth=0.1) + geom_vline(xintercept = 0) + xlim(c(-20,20))
show(plot)

```



::::::

## Simulating multiple sampling distributions

- To create multiple sampling distributions, we need to use another `for loop` to repeat the process of creating sampling distributions. There are 2 things to add.
    1. Repeat each simulation by wrapping our code in another `for loop`
    2. Save the precisions from each sampling distribution

## 2. Sampling distributions for multiple sample sizes

:::::: panel-tabset

## Code Puzzle

::::: code-puzzle

::: static

AllSampleSizes <- c(9, 25, 49, 81, 121)
SD <- 12
AllPrecisions <- {{c()}}

:::

::: indent
for ({{SampleSize in AllSampleSizes}}){

   
    # Code from first simulation, starting from line 4 ...

    AllPrecisions <- {{c(AllPrecisions, summary_table$Precision)}}

}
:::

:::::

<!--

## Instructions

::: {.nonincremental .normal-list}

- [ ] Define some variables

    ```r
    AllSampleSizes <- c(9, 25, 49, 81, 121)
    SD <- 12
    AllPrecisions <- c() # Empty list to store all precision calculations
    ```

- [ ] Repeat the simulation: 
    - [ ] Create a new `for loop`

        ```r
        for (SampleSize in AllSampleSizes){
            ...
        }
        ```

    - [ ] Copy the code from the `Sampling distribution simulation` slide into the `for loop`
- [ ] To store the simulated precisions, add the following to the end of the `for loop`

    ```r
    AllPrecisions <- c(AllPrecisions, summary_table$Precision)
    ```

:::

-->

## Plots

```{r}

AllSampleSizes <- c(9, 25, 49, 81, 121)
SD <- 12
AllPrecisions <- c()
for (SampleSize in AllSampleSizes){
    AllExperiments <- 1:5000
simulated_means <- c()
for (experiment in AllExperiments){
    simulated_data <- rnorm(n = SampleSize, mean = Mean, sd = SD)
    simulated_means[experiment] <- mean(simulated_data)
}
sampling_distribution <- tibble(Means=simulated_means)
summary_table <- sampling_distribution %>%
summarize(SE = sd(Means), Precision = 1/SE)
summary_table
plot <- ggplot(sampling_distribution, aes(x=Means)) +
geom_histogram(binwidth=0.1) + geom_vline(xintercept = 0) + xlim(c(-20,20))
show(plot)
    AllPrecisions <- c(AllPrecisions, summary_table$Precision)
}

```


::::::

## 4. Plot precision vs. sample size

:::::: panel-tabset

## Code Puzzle

::::: code-puzzle
::: free
precision_v_samplesize <- tibble(
    SampleSize={{{AllSampleSizes}}}, 
    Precision={{{AllPrecisions}}})
:::

::: free
ggplot({{precision_v_samplesize}}, aes(x={{SampleSize}}, y={{Precision}})) + 
    {{geom_point() + geom_line()}}
:::

:::::

## Plot

```{r}


precision_v_samplesize <- tibble(
SampleSize=AllSampleSizes,
Precision=AllPrecisions)
ggplot(precision_v_samplesize, aes(x=SampleSize, y=Precision)) +
geom_point() + geom_line()


```

::::::

## 5. Simulate sampling distributions for multiple standard deviations

:::::: panel-tabset

## Code Puzzle

Same as above, we need to repeat the simulation, but this time for multiple SDs

::::: code-puzzle

::: static
SampleSize <- 49
AllSDs <- c(4, 8, 12, 16, 20)
AllPrecisions <- {{{c()}}}
:::

::: indent
for ({{SD}} in {{AllSDs}}){
    # Code from first simulation, starting from line 4
    {{{AllPrecisions}}} <- c(AllPrecisions, summary_table$Precision)
}
:::

:::::

## Plots

```{r}

SampleSize <- 49
AllSDs <- c(4, 8, 12, 16, 20)
AllPrecisions <- c()
for (SD in AllSDs){
    AllExperiments <- 1:5000
simulated_means <- c()
for (experiment in AllExperiments){
    simulated_data <- rnorm(n = SampleSize, mean = Mean, sd = SD)
    simulated_means[experiment] <- mean(simulated_data)
}
sampling_distribution <- tibble(Means=simulated_means)
summary_table <- sampling_distribution %>%
summarize(SE = sd(Means), Precision = 1/SE)
summary_table
plot <- ggplot(sampling_distribution, aes(x=Means)) +
geom_histogram(binwidth=0.1) + geom_vline(xintercept = 0) + xlim(c(-20,20))
show(plot)
    AllPrecisions <- c(AllPrecisions, summary_table$Precision)
}

```


::::::

## 6. Plot precision vs. sample standard deviation

:::::: panel-tabset

## Code Puzzle

::::: code-puzzle

::: free
precision_v_sd <- tibble(
    SD={{{AllSDs}}}, 
    Precision={{{AllPrecisions}}})

:::

::: free
ggplot({{precision_v_sd}}, aes(x={{SD}}, y={{Precision}})) + 
    {{geom_point() + geom_line()}}

:::

:::::

## Plot

```{r}


precision_v_sd <- tibble(
SD=AllSDs,
Precision=AllPrecisions)
ggplot(precision_v_sd, aes(x=SD, y=Precision)) +
geom_point() + geom_line()


```

::::::

## Deliverables checklist

::: {.normal-list .nonincremental}
- [ ] Sampling distribution plot with sample size = 49; SD = 12
- [ ] Sampling distribution plot with sample size between 9, 25, 49, 81, 121; SD = 12
- [ ] Plot of precision vs. sample size
- [ ] Sampling distribution plot with sample size = 49; vary SD between 4, 8, 12, 16, 20
- [ ] Plot of precision vs. sample standard deviation
- [ ] Answers to the following questions:
    - [ ] What is the relationship between sample size and precision? Why?
    - [ ] What is the relationship between SD and precision? Why?
        - Hint: Higher SD $\rightarrow$ Higher/Lower __ $\rightarrow$ Higher/Lower __ $\rightarrow$ Higher/Lower Precision
:::