---
title: "9: Effect sizes"
---

## Key question


- Sometimes, there is a true difference means, but the difference is too small to be meaningful.
- How do we know that a difference is big enough to meaningful? 
    - Answer: Effect sizes.

## Effect sizes

- The effect size is how many standard deviations the means are apart. 
    - An effect size of 1 means that the differences in means are 1 SD apart.
    - Basically: the SD becomes the new unit
- It is a measure of signal strength, so it does not depend on sample size.
- To calculate, use: 
    - (mean(x1) - mean(x2)) / sd(x2)
- You can get the means and SD from from your summary table like this (make sure to replace the names with the appropriate values)

    ```r
    summary_table[row_number,column_number]

    ```

## What we will do

::: card-list

1. Today, we will be seeing if there is a meaningful difference in annual salary between UT grads and Texas A&M grads. (Note: these are simulated, so fake, data!)
1. We will look at a data set with a [large sample size](/data/salsLargeN.csv){.filelink}, and separately, adataset with a [small sample size](/data/salsSmallN.csv){.filelink}. Both are complete datasets.
1. Using effect sizes, we will assess if a difference in means is meaningful, despite a small p-value.
    
:::

## Review

Log in and answer directly in the slides!


::: iframe

https://app.sli.do/event/i3EEgRwm48fNt3yUXKL7ys/embed/polls/d570183d-8d17-4870-94bb-1614fa6c6378

:::


## Slido

Here is a sneak peak of a dataset:

```{r}

library(tidyverse)
library(ggplot2)

data1 <- read_csv("./data/salsLargeN.csv")
data2 <- read_csv("./data/salsSmallN.csv")

head(data1)

```

Based on the above, what is the code to conduct a t.test?

::: iframe

https://app.sli.do/event/i3EEgRwm48fNt3yUXKL7ys/embed/polls/3041c270-0d66-4873-83f8-12b864440efc

:::



# Deliverables 

Multiple slides, make sure to go to the end!

## {.center}

- For each dataset, create a frequency polygon, CI plot, summary table, and t-test, grouping by `school`

# What your results should look like

## Large sample  

```{r}

library(tidyverse)
library(ggplot2)

data <- data1

hist <- ggplot(data, aes(x=salary, color=school)) + geom_freqpoly()

t_test <- t.test(data=data, salary ~ school)

summary <- data %>% group_by(school) %>% summarize(SampleSize=n(), Mean = mean(salary), SD = sd(salary), SE=SD/sqrt(SampleSize), LowerCI = Mean-1.96*SE, UpperCI = Mean+1.96*SE)


mean_plot <- ggplot(summary, aes(x=school, y=Mean)) + geom_point() + geom_errorbar(aes(ymin=LowerCI, ymax=UpperCI))  

```

:::: panel-tabset

## Frequency polygon

```{r}
hist
```

## Summary table
```{r}

knitr::kable(summary)


```

## Mean plot

```{r}
mean_plot
```

## t-test

```{r}

print(t_test)

```

::::

## Small sample

```{r}
data <- data2

hist <- ggplot(data, aes(x=salary, color=school)) + geom_freqpoly(binwidth=1000)

t_test<- t.test(data=data, salary ~ school)

summary <- data %>% group_by(school) %>% summarize(SampleSize=n(), Mean = mean(salary), SD = sd(salary), SE=SD/sqrt(SampleSize), LowerCI = Mean-1.96*SE, UpperCI = Mean+1.96*SE)

mean_plot <- ggplot(summary, aes(x=school, y=Mean)) + geom_point() + geom_errorbar(aes(ymin=LowerCI, ymax=UpperCI))

```

:::: panel-tabset

## Frequency polygon

```{r}
hist
```

## Summary table
```{r}

knitr::kable(summary)


```

## Mean plot

```{r}
mean_plot
```

## t-test

```{r}

print(t_test)

```

::::


## For the large dataset, answer the following

::: card-list
- Are the frequency polygons similar or different?
- Does the confidence interval include zero? Or is it above/below?
- Is the p-value "small" or "big"?
- Visually compare the *sample* means between schools in the frequency polygon (not population!). Is this difference "small" or "big"?
- Compare this difference in means to the p-value/CI. What is "weird" about these results? Where there are "/", choose the right words for following

    ```{verbatim}
    The p-value is "small/big" and the sample mean difference is "small/big". 
    The p-value shows "strong/weak" evidence for a difference, 
    and the frequency polygon shows "strong/weak" evidence for a difference. 
    So p-value and frequency polygon are "consisistent/inconsistent" with each other.
    ```

:::

## For the small dataset, answer the following

- Are the frequency polygons similar or different?
- Does the confidence interval include zero? Or is it above/below?
- Is the p-value "small" or "big"?
- Visually compare the *sample* means between schools in the frequency polygon (not population!). Is this difference "small" or "big"?
- Compare this difference in means to the p-value/CI. What is "weird" about these results? Where there are "/", choose the right words for following

    ```{verbatim}
    The p-value is "small/big" and the sample mean difference is "small/big". 
    The p-value shows "strong/weak" evidence for a difference, 
    and the frequency polygon shows "strong/weak" evidence for a difference. 
    So p-value and frequency polygon are "consisistent/inconsistent" with each other.
    ```

<!--

## What's weird?

::: card-list

- For the large dataset
    - The p-value is very small, this means that the probability of getting this difference, given the null is true, is small. This provides evidence that there is a true difference.
    - But the frequency polygons look very similar!
    - Key question: So is this difference big enough to be an important difference?
- For the small dataset
    - The p-value is larger, so we cannot reject the null hypothesis.
    - The frequency polygons look different, but since it's a small sample, it could be due to noise.

- When our p-value/CI suggest a true difference, how can we determine if that difference is big enough to be meaningful?

:::

-->


##

- Calculate the effect size for the large dataset. 
- What is the effect size, and in your opinion, is this big or small? (Think about how effect size relate to standard deviation)
- What is your main takeaway from today's lab?





